---
title: "Cell annotation and visualization"
subtitle: ''
author: "Shahin Mohammadi"
date: "Run on `r Sys.time()`"
documentclass: article
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(ACTIONet)
library(data.table)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(corrplot)
library(muscat)
library(synapser)
library(synExtra)
synLogin(rememberMe = TRUE)
source("functions.R")
})

knitr::opts_chunk$set(
	eval = FALSE,
	error = FALSE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	dev = c("png", "pdf"),
	include = FALSE,
	tidy = FALSE
)


```


```{r}
# devtools::load_all("~/ACTIONet/")
install.packages("~/ACTIONet/", repos = NULL)

```


```{r}
dataset.path = "~/results/datasets/"
figures.path = "~/results/figures"
tables.path = "~/results/tables"
input.path = "~/results/input"

colors = loadDataset("celltype_colors", dataset.path = dataset.path)
ACTIONet_summary = loadDataset("ACTIONet_summary", dataset.path = dataset.path)

```




# Plot main ACTIONet
## Celltypes
```{r}
gg = plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$Labels, palette = colors, text_size = 2, use_repel = T)

storeFigure(gg, name = "ACTIONet_annotated", extension = "png", figures.path = figures.path, width = 8, height = 6)

```

## Supps
### Phenotype
```{r}
gg = plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$Phenotype, palette = c("#cccccc", "#888888"), text_size = 2, use_repel = T, add_text_labels = F, point_size = 0.5)

storeFigure(gg, name = "ACTIONet_phenotype_annotated_no_labels", extension = "png", figures.path = figures.path, width = 8, height = 6)

```

### Batch
```{r}
gg = plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$Batch, text_size = 2, use_repel = T, add_text_labels = F, point_size = 0.5)

storeFigure(gg, name = "ACTIONet_batch_annotated_no_labels", extension = "png", figures.path = figures.path, width = 8, height = 6)

```

### Gender
```{r}
mask = !is.na(ACTIONet_summary$metadata$Gender)
gg = plot.ACTIONet(ACTIONet_summary$ACTIONet2D[mask, ], ACTIONet_summary$metadata$Gender[mask], text_size = 2, use_repel = T, add_text_labels = F, point_size = 0.5, palette = c("pink", "#91bfdb"))

storeFigure(gg, name = "ACTIONet_gender_annotated_no_labels", extension = "png", figures.path = figures.path, width = 8, height = 6)

```

### Archetype
```{r}
gg = plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$assigned_archetype, text_size = 2, use_repel = T, add_text_labels = F, point_size = 0.5, palette = as.character(pals::polychrome(31)))

storeFigure(gg, name = "ACTIONet_archetypes_annotated_no_labels", extension = "png", figures.path = figures.path, width = 8, height = 6)

```

### Dataset
```{r}
gg = plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$dataset, text_size = 2, use_repel = T, add_text_labels = F, point_size = 0.5, palette = c("#f1a340", "#998ec3"))


storeFigure(gg, name = "ACTIONet_datasets_annotated_no_labels", extension = "png", figures.path = figures.path, width = 8, height = 6)

```


## Plot ACTIONet plots per dataset
### McLean dataset
```{r}
mask = ACTIONet_summary$metadata$dataset == "McLean"
DS1.coors = ACTIONet_summary$ACTIONet2D[mask, ]
DS1.labels = ACTIONet_summary$metadata$Labels[mask]

gg = plot.ACTIONet(DS1.coors, DS1.labels, palette = colors, text_size = 2, use_repel = T)

storeFigure(gg, name = "ACTIONet_annotated_DS1", extension = "png", figures.path = figures.path, width = 8, height = 6)

```


### MtSinai dataset
```{r}
mask = ACTIONet_summary$metadata$dataset == "MtSinai"
DS2.coors = ACTIONet_summary$ACTIONet2D[mask, ]
DS2.labels = ACTIONet_summary$metadata$Labels[mask]

gg = plot.ACTIONet(DS1.coors, DS1.labels, palette = colors, text_size = 2, use_repel = T)

storeFigure(gg, name = "ACTIONet_annotated_DS2", extension = "png", figures.path = figures.path, width = 8, height = 6)

```


# Plot cell type fraction stats
## Identify and filter outlier samples (based on the % of Ex cells)
In total, 5 samples are filtered, all of which are SZ (SZ33, SZ24, SZ29, SZ3, and SZ15)

```{r}
require(ggpubr)

ncells = apply(table(ACTIONet_summary$metadata$Labels, ACTIONet_summary$metadata$Individual), 2, as.numeric)
rownames(ncells) = levels(ACTIONet_summary$metadata$Labels)

cs = Matrix::colSums(ncells)
ncells.freq = 100*scale(ncells, center = F, scale = cs)
Ex.perc = (fast_column_sums(ncells.freq[grepl("^Ex", rownames(ncells.freq)), ]))

df = data.frame(sample = colnames(ncells), perc = Ex.perc)
gg = ggdensity(df, x = "perc", fill = "lightgray",
   add = "mean", rug = TRUE) + xlim(c(0, 100))

storeFigure(gg, name = "Ex_perc_density", extension = "png", figures.path = figures.path, width = 8, height = 6)

ncells.freq = ncells.freq[, order(Ex.perc, decreasing = T)]
Ex.perc = (fast_column_sums(ncells.freq[grepl("^Ex", rownames(ncells.freq)) & !grepl("^Ex-NRGN", rownames(ncells.freq)), ]))

mask = (Ex.perc >= 10) & (Ex.perc <= 80) 
filtered.samples = colnames(ncells.freq)[which(!mask)]
ncells.freq = ncells.freq[, mask]


```


## Plot cell type fractions for the rest of samples
```{r}
X = ncells.freq
df = reshape2::melt(X)
colnames(df)=c("celltype", "sample", "freq")

df$celltype = factor(df$celltype, names(colors))

gg = ggbarplot(df, "sample", "freq",
  fill = "celltype", color = "black", palette = colors[levels(df$celltype)], xlab = "Individual", ylab = "Percentage") + theme(axis.text.x = element_text(face="bold", size=8, angle=90,hjust=0.95,vjust=0.2), axis.text.y = element_text(face="bold", size=12, angle=0), axis.title = element_text(face="bold", size=18, angle=0))

storeFigure(gg, name = "celltype_perc_joint", extension = "png", figures.path = figures.path, width = 8, height = 6)

```

## Plot cell type fraction stats per phenotype
```{r}
n_cells = table(ACTIONet_summary$metadata$Labels, ACTIONet_summary$metadata$Individual)

X.CON = apply(n_cells[, grep("CON", colnames(n_cells))], 2, as.numeric)
X.SZ = apply(n_cells[, grep("SZ", colnames(n_cells))], 2, as.numeric)
rownames(X.CON) = rownames(X.SZ) = levels(ACTIONet_summary$metadata$Labels)


Perc = t(apply(cbind(Matrix::rowMeans(X.CON), Matrix::rowMeans(X.SZ)), 1, function(x) 100*x / sum(x)))
colnames(Perc) = c("CON", "SZ")
Perc = Perc[order(Perc[, 1] - Perc[, 2], decreasing = T), ]

df = reshape2::melt(Perc)
colnames(df) = c("Celltype", "Phenotype", "Perc")
df$Celltype = factor(df$Celltype, names(colors))

gg = ggbarplot(df, "Celltype", "Perc", fill = "Phenotype", palette = c("#cccccc", "#888888"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, colour = colors)) + xlab("Celltype") + ylab("Percentage")

storeFigure(gg, name = "Celltype_perc_per_phenotype", extension = "png", figures.path = figures.path, width = 8, height = 6)

```



# Plot gene/umi statistics
## Per cell type
```{r}
umis = ACTIONet_summary$metadata$umi_count
mito.perc = ACTIONet_summary$metadata$mito_perc
genes = ACTIONet_summary$metadata$gene_counts
dataset = ACTIONet_summary$metadata$dataset
indiv = ACTIONet_summary$metadata$Individual
celltype = ACTIONet_summary$metadata$Labels
df = data.frame(celltype = celltype, umis = umis, genes = genes, mito = mito.perc, dataset = dataset, individual = indiv)  

df$celltype = factor(df$celltype, names(colors))

```

### McLean
```{r}
require(ggpubr)
gg = ggviolin(df[df$dataset == "McLean", ], "celltype", "umis", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("UMIs")+ theme(legend.position = "none")
storeFigure(gg, name = "McLean_umis", extension = "png", figures.path = figures.path, width = 8, height = 6)


require(ggpubr)
gg = ggviolin(df[df$dataset == "McLean", ], "celltype", "genes", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("genes")+ theme(legend.position = "none")
storeFigure(gg, name = "McLean_genes", extension = "png", figures.path = figures.path, width = 8, height = 6)


gg = ggviolin(df[df$dataset == "McLean", ], "celltype", "mito", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("mito") + ylim(c(0, 10))+ theme(legend.position = "none")
storeFigure(gg, name = "McLean_mito", extension = "png", figures.path = figures.path, width = 8, height = 6)

```

### MtSinai
```{r}
gg = ggviolin(df[df$dataset == "MtSinai", ], "celltype", "umis", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("UMIs")+ theme(legend.position = "none")
storeFigure(gg, name = "MtSinai_umis", extension = "png", figures.path = figures.path, width = 8, height = 6)


gg = ggviolin(df[df$dataset == "MtSinai", ], "celltype", "genes", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("genes")+ theme(legend.position = "none")
storeFigure(gg, name = "MtSinai_genes", extension = "png", figures.path = figures.path, width = 8, height = 6)


gg = ggviolin(df[df$dataset == "MtSinai", ], "celltype", "mito", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("mito") + ylim(c(0, 10))+ theme(legend.position = "none")
storeFigure(gg, name = "MtSinai_mito", extension = "png", figures.path = figures.path, width = 8, height = 6)


```

## Per sample
```{r}
umis.per.ind = split(ACTIONet_summary$metadata$umi_count, ACTIONet_summary$metadata$Individual)
mean.umis = sapply(umis.per.ind, mean)

df$individual = factor(as.character(df$individual), names(mean.umis)[order(mean.umis, decreasing = T)])

```

### McLean
```{r}
require(ggpubr)
gg = ggviolin(df[df$dataset == "McLean", ], "individual", "umis", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("individual") + ylab("UMIs")+ theme(legend.position = "none")
storeFigure(gg, name = "McLean_umis_per_sample", extension = "png", figures.path = figures.path, width = 11, height = 6)


require(ggpubr)
gg = ggviolin(df[df$dataset == "McLean", ], "individual", "genes", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("individual") + ylab("genes")+ theme(legend.position = "none")
storeFigure(gg, name = "McLean_genes_per_sample", extension = "png", figures.path = figures.path, width = 11, height = 6)


gg = ggviolin(df[df$dataset == "McLean", ], "individual", "mito", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("individual") + ylab("mito") + ylim(c(0, 10))+ theme(legend.position = "none")
storeFigure(gg, name = "McLean_mito_per_sample", extension = "png", figures.path = figures.path, width = 11, height = 6)

```

### MtSinai
```{r}
gg = ggviolin(df[df$dataset == "MtSinai", ], "individual", "umis", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("individual") + ylab("UMIs")+ theme(legend.position = "none")
storeFigure(gg, name = "MtSinai_umis_per_sample", extension = "png", figures.path = figures.path, width = 18, height = 6)


gg = ggviolin(df[df$dataset == "MtSinai", ], "individual", "genes", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("individual") + ylab("genes")+ theme(legend.position = "none")
storeFigure(gg, name = "MtSinai_genes_per_sample", extension = "png", figures.path = figures.path, width = 18, height = 6)


gg = ggviolin(df[df$dataset == "MtSinai", ], "individual", "mito", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("individual") + ylab("mito") + ylim(c(0, 10))+ theme(legend.position = "none")
storeFigure(gg, name = "MtSinai_mito_per_sample", extension = "png", figures.path = figures.path, width = 18, height = 6)


```







# Plot mappings of cell type annotations to other annotations
## Load markers
```{r}
data("curatedMarkers_human")
Layer.markers = loadInputDataset("Maynard_layer_marker", "rds")
Layer.markers = Layer.markers[c(2:7, 1)]
Velmeshev.markers = loadInputDataset("Velmeshev", "tsv")
Velmeshev.markers = split(Velmeshev.markers$`Gene name`, Velmeshev.markers$`Cell type`)
Mohammadi.markers = curatedMarkers_human$Brain$PFC$Mohammadi2020$marker.genes

celltype.gene.spec = loadDataset("celltype_gene_specificity")

rdbu_fun = circlize::colorRamp2(c(-3, -1, 0, 1, 3), rev(pals::brewer.rdbu(9)[seq(1, 9, by = 2)]))

```

## Annotate
### Layers (Maynard et al., 2021)
```{r}
X = as(do.call(cbind, lapply(Layer.markers, function(gs) as.numeric(rownames(celltype.gene.spec) %in% gs))), "dgCMatrix")
rownames(X) = rownames(celltype.gene.spec)
Layer.annot = assess.geneset.enrichment.from.scores(celltype.gene.spec, X)

Z  = scale(t(Layer.annot$logPvals))
rownames(Z) = colnames(celltype.gene.spec)
colnames(Z) = c(paste("L", c(1:6), sep = ""), "WM")

ht = Heatmap(Z, cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Layer", column_title = "Layer (Maynard et al.)")

storeFigure(ht, name = "Layer_annotation_Maynard", extension = "pdf", figures.path = figures.path, width = 4, height = 8)


```

### Cell types (Mohammadi et al., 2019)
```{r}
X = as(do.call(cbind, lapply(curatedMarkers_human$Brain$PFC$Mohammadi2020$marker.genes, function(gs) as.numeric(rownames(celltype.gene.spec) %in% gs))), "dgCMatrix")
rownames(X) = rownames(celltype.gene.spec)
Celltype.annot = assess.geneset.enrichment.from.scores(celltype.gene.spec, X)

Z  = scale(t(Celltype.annot$logPvals))
rownames(Z) = colnames(celltype.gene.spec)

M = as(t(MWM_hungarian(t(Celltype.annot$logPvals))), "dgTMatrix")

ht = Heatmap(Z[, M@i+1], cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Celltypes", column_title = "Celltypes (Mohammadi et al.)")

storeFigure(ht, name = "celltype_annotation_mohammadi_markers", extension = "pdf", figures.path = figures.path, width = 6, height = 8)

```

### Cell types (Velmeshev et al., 2019)
```{r}
X = as(do.call(cbind, lapply(curatedMarkers_human$Brain$PFC$Velmeshev2019$marker.genes, function(gs) as.numeric(rownames(celltype.gene.spec) %in% gs))), "dgCMatrix")
rownames(X) = rownames(celltype.gene.spec)
Celltype.annot = assess.geneset.enrichment.from.scores(celltype.gene.spec, X)


Z  = scale(t(Celltype.annot$logPvals))
rownames(Z) = colnames(celltype.gene.spec)

M = as(t(MWM_hungarian(t(Celltype.annot$logPvals))), "dgTMatrix")


ht = Heatmap(Z[, M@i+1], cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Celltypes", column_title = "Celltypes (Velmeshev et al.)")

storeFigure(ht, name = "celltype_annotation_velmeshev_markers", extension = "pdf", figures.path = figures.path, width = 6, height = 8)

```
### Cell types (Mathys et al., 2019)
```{r}
X = as(do.call(cbind, lapply(curatedMarkers_human$Brain$PFC$MathysDavila2019$marker.genes, function(gs) as.numeric(rownames(celltype.gene.spec) %in% gs))), "dgCMatrix")
rownames(X) = rownames(celltype.gene.spec)
Celltype.annot = assess.geneset.enrichment.from.scores(celltype.gene.spec, X)



Z  = scale(t(Celltype.annot$logPvals))
rownames(Z) = colnames(celltype.gene.spec)


M = as(t(MWM_hungarian(t(Celltype.annot$logPvals))), "dgTMatrix")

ht = Heatmap(Z[, M@i+1], cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Celltypes", column_title = "Celltypes (Mathys and Davilla et al.)")

storeFigure(ht, name = "celltype_annotation_mathys_markers", extension = "pdf", figures.path = figures.path, width = 10, height = 8)

```

## Bulk Layers (He et al., 2017)
```{r}
X = as(do.call(cbind, lapply(curatedMarkers_human$Brain$Layers$marker.genes, function(gs) as.numeric(rownames(celltype.gene.spec) %in% gs))), "dgCMatrix")
rownames(X) = rownames(celltype.gene.spec)
Celltype.annot = assess.geneset.enrichment.from.scores(celltype.gene.spec, X)



Z  = scale(t(Celltype.annot$logPvals))
rownames(Z) = colnames(celltype.gene.spec)

ht = Heatmap(Z, cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Layers", column_title = "Layers (He et al.)")

storeFigure(ht, name = "Layer_annotation_He_markers", extension = "pdf", figures.path = figures.path, width = 4, height = 8)

```









