---
title: "Analyze DE genes"
subtitle: "Step 3: Postprocessing the alignment of DE results across datasets"

output: html_notebook
---

# Setup
```{r include=FALSE}
require(ACTIONet)
require(stringr)
require(ComplexHeatmap)


results.path = "~/results"
input.path = "~/results/input"
dataset.path = "~/results/datasets"
tables.path = "~/results/tables"
figures.path = "~/results/figures"

```


## Enrichment function
```{r}
assess.genesets <-function (arch.gs, terms.gs, N, min.pval = 1e-100, correct = "none"){
    shared = t(sapply(terms.gs, function(gs1) {
        sapply(arch.gs, function(gs2) {
            nn = intersect(gs1, gs2)
        })
    }))
    colnames(shared) = names(arch.gs)
    GS.sizes = sapply(terms.gs, length)
    logPvals.out = sapply(1:ncol(shared), function(i) {
        gg = shared[, i]
        x = as.numeric(sapply(gg, length))
        n.sample = length(arch.gs[[i]])
        n.success = as.numeric(GS.sizes)
        v = rep(1, length(x))
        min.overlap = n.success * n.sample/N
        idx = which(x >= min.overlap)
        if (length(idx) == 0) 
            return(v)
        v[idx] = (phyper(x[idx]-1, n.sample, N-n.sample, n.success[idx], lower.tail = F))
        # v[idx] = HGT_tail(population.size = N, success.count = n.success[idx], 
        #     sample.size = n.sample, observed.success = x[idx])
        return(v)
    })
    if(correct == "global") {
      logPvals.out = matrix(p.adjust(logPvals.out, "fdr"), nrow = nrow(logPvals.out))
    } else if(correct == "local") {
      logPvals.out = apply(logPvals.out, 2, function(x) p.adjust(x, "fdr"))
    }
    rownames(logPvals.out) = names(terms.gs)
    colnames(logPvals.out) = names(arch.gs)
    return(-log10(Matrix::t(logPvals.out)))
}
```

# Load primary datasets
```{r, eval = T}
ACTIONet_summary = readr::read_rds(file.path(dataset.path, "ACTIONet_summary_filtered_individuals.rds"))
pb.logcounts = readr::read_rds(file.path(dataset.path, "PB_mean_logcounts_final.RDS"))

color.df = readRDS(file.path(dataset.path, "celltype_colors.rds"))
colors = color.df$color
names(colors) = color.df$celltype

```

# Load DE results
```{r, eval = T}
resDE = readr::read_rds(file.path(dataset.path, "Cohort_specific_DE_results.rds"))
filtered.tables = readr::read_rds(file.path(dataset.path, "Cohort_specific_DE_results_filtered.rds"))
combined.analysis.tables = readr::read_rds(file.path(dataset.path, "meta_analysis_results.rds"))

DE.new = readRDS(file.path(dataset.path, "DE_genes_pseudobulk.rds"))
Up.genes = DE.new$Up.genes
Down.genes = DE.new$Down.genes
DE.sc = DE.new$DE.sc
ordered.celltypes = rownames(X)[order(apply(X, 1, sum), decreasing = T)]

```



# Compute overlap of selected DE genes with bulk (PEC)
```{r}
SZ.genes = readr::read_rds(file.path(input.folder, "SCZ_associated_genesets.rds"))

Up.genes.overlap = sapply(Up.genes, function(gs) intersect(gs, SZ.genes$`DE.Up (PEC)`))
Up.genes.size = sapply(Up.genes.overlap, length) 
Up.En = assess.genesets(Up.genes[ordered.celltypes], SZ.genes[c(5, 6)], nrow(pb.logcounts), correct = "local")[, 1]

Down.genes.overlap = sapply(Down.genes, function(gs) intersect(gs, SZ.genes$`DE.Down (PEC)`))
Down.genes.size = sapply(Down.genes.overlap, length) 
Down.En = assess.genesets(Down.genes[ordered.celltypes], SZ.genes[c(5, 6)], nrow(pb.logcounts), correct = "local")[, 2]


DE.overlap.tbl = data.frame(Celltype = ordered.celltypes, Up = sapply(Up.genes[ordered.celltypes], length), Up_vs_bulk_count = Up.genes.size, Up_vs_bulk_enrichment = Up.En, Down = sapply(Down.genes[ordered.celltypes], length), Down_vs_bulk_count = Down.genes.size, Down_vs_bulk_enrichment = Down.En)

write.table(DE.overlap.tbl, file.path(tables.path, "DE_vs_bulk_overlap.tsv"), sep = "\t", row.names = F, col.names = T, quote = F)
```


# Plot the total number of DE genes
```{r}
celltype.colors = colors[names(Up.genes)]
# names(celltype.colors) = names(Up.genes)

df.Up = data.frame(Counts = sapply(Up.genes, function(x) length(setdiff(x, SZ.genes$`DE.Up (PEC)`))), Celltype = names(Up.genes), Direction="Up", Color = celltype.colors[names(Up.genes)], stringsAsFactors = F)

df.UpandBulk = data.frame(Counts = sapply(Up.genes, function(x) length(intersect(x, SZ.genes$`DE.Up (PEC)`))), Celltype = names(Up.genes), Direction="Up & Bulk", Color = celltype.colors[names(Up.genes)], stringsAsFactors = F)


df.Down = data.frame(Counts = -sapply(Down.genes, function(x) length(setdiff(x, SZ.genes$`DE.Down (PEC)`))), Celltype = names(Down.genes), Direction="Down", Color = celltype.colors[names(Down.genes)], stringsAsFactors = F)

df.DownandBulk = data.frame(Counts = -sapply(Down.genes, function(x) length(intersect(x, SZ.genes$`DE.Down (PEC)`))), Celltype = names(Down.genes), Direction="Down & Bulk", Color = celltype.colors[names(Down.genes)], stringsAsFactors = F)



df = rbind(df.Up, df.UpandBulk, df.Down, df.DownandBulk)

total.Up = sapply(Up.genes, length)
total.Down = sapply(Down.genes, length)

set.seed(0)
# total = total.Up + total.Down + 0.001*rnorm(length(Up.genes))
total = apply(cbind(total.Up, total.Down), 1, max)

arch.perm = order(total, decreasing = F)
df$Celltype = factor(df$Celltype, rev(ordered.celltypes))

pdf(file.path(figures.path, "NumDysregGenes.pdf"), width = 8, height = 5)
ggplot(data = df, aes(x = Celltype, y = Counts, fill = Direction)) + geom_bar(stat = "identity")+
  coord_flip()+ylab("Sorted Celltypes")+
labs(y = "# Genes", x = "Sorted Celltypes")+
  theme_minimal()+
  guides(fill = FALSE)+ scale_fill_manual(values=c("#3288bd", colorspace::darken("#3288bd", 0.35), "#d53e4f", colorspace::darken("#d53e4f", 0.35))) + theme(axis.text.y = element_text(face="bold", color=celltype.colors[levels(df$Celltype)], size=12, angle=0), axis.text.x = element_text(face="bold", color="black", size=12, angle=0), axis.title = element_text(face="bold", size=14, angle=0))
dev.off()

```


# Visualize selected genes
## Read bulk DE results
```{r}
PEC.DE = read.table(file.path(input.path, "PEC_DE_table.csv"), header = T)
common.genes = intersect(rownames(pb.logcounts), PEC.DE$gene_name)
PEC.tstat = PEC.DE$SCZ.t.value[match(common.genes, PEC.DE$gene_name)]
names(PEC.tstat) = common.genes

```

## Plot heatmap
```{r}
selected.genes.tbl = read.xlsx(file.path(dataset.path, "Top50Genes.xlsx"))
selected.genes = sort(unique(unlist(as.list(selected.genes.tbl))))
selected.genes = intersect(selected.genes, names(PEC.tstat))

X = DE.new$DE.sc[selected.genes, ]
perm = order(fast_row_sums(X), decreasing = T)
Y = as.matrix(PEC.tstat[selected.genes])
colnames(Y) = c("Bulk")
RdBu.pal = circlize::colorRamp2(seq(-3, 3, length.out = 7), rev(pals::brewer.rdbu(7)))

pdf(file.path(figures.path, "Top50_selected_DE_genes.pdf"), width = 6, height = 16)
Heatmap(X[perm, ], cluster_rows = F, rect_gp = gpar(col = "black"), cluster_columns = F, column_title = "Celltypes", column_title_gp = gpar(fontsize = 21), column_names_gp = gpar(col = colors[colnames(DE.new$DE.sc)]), name = "DE", col = RdBu.pal, row_names_side = "left") + Heatmap(Y[perm, ], cluster_rows = F, rect_gp = gpar(col = "black"), cluster_columns = F, name = "Bulk", col = RdBu.pal, row_names_side = "left")
dev.off()

```


# Functional enrichment
## Load annotations
```{r}
FunCat = readRDS(file.path(dataset.path, "FunCat.rds"))

FunCat.genes = split(FunCat$FunCat2Gene$Gene, factor(FunCat$FunCat2Gene$Category, unique(FunCat$FunCat2Gene$Category)))[-15]
names(FunCat.genes) = FunCat$FunCat2Class$Category

FunCat.annotation = FunCat$FunCat2Class$Classification

FunCatPal = ggpubr::get_palette("npg", length(unique(FunCat.annotation)))
names(FunCatPal) = unique(FunCat.annotation)


```

## Perform enrichment
```{r}
Up.enrichment = assess.genesets(FunCat.genes, Up.genes[1:17], nrow(pb.logcounts), correct = "local")
Down.enrichment = assess.genesets(FunCat.genes, Down.genes[1:17], nrow(pb.logcounts), correct = "local")

Up.enrichment[Up.enrichment < -log10(0.05)] = 0
Down.enrichment[Down.enrichment < -log10(0.05)] = 0

ha_rows = rowAnnotation(df = list("Class" = FunCat.annotation), col = list("Class" = FunCatPal), annotation_legend_param = list("Class"=list(title_gp = gpar(fontsize = 0), labels_gp = gpar(fontsize = 10))))


X.U = (Up.enrichment)
# redCol_fun = circlize::colorRamp2(c(quantile(X.U, 0.25), quantile(X.U, 0.5), quantile(X.U, 0.85), quantile(X.U, 0.99)), c("#ffffff", "#fee5d9", "#ef3b2c", "#99000d"))
redCol_fun = circlize::colorRamp2(c(0, exp(quantile(log(X.U)[X.U > -log10(0.05)], seq(0.05, 0.95, length.out = 12)))), c("#ffffff", pals::brewer.reds(12)))

X.D = (Down.enrichment)
# blueCol_fun = circlize::colorRamp2(c(quantile(X.D, 0.25), quantile(X.U, 0.5), quantile(X.D, 0.85), quantile(X.D, 0.99)), c( "#ffffff", "#9ecae1", "#2171b5", "#08306b"))

blueCol_fun = circlize::colorRamp2(c(0, exp(quantile(log(X.D)[X.D > -log10(0.05)], seq(0.05, 0.95, length.out = 12)))), c("#ffffff", pals::brewer.blues(12)))

# pals::brewer.reds()


pdf(file.path(figures.path, "DE_FunCat.pdf"), width = 14, height = 7)
par(mar=c(0,150,0,0))
Heatmap(X.U, rect_gp = gpar(col = "black"), name = "Up", column_title = "Up-regulated", cluster_rows = F, cluster_columns = F, col = redCol_fun, row_names_side = "left",  column_names_gp = gpar(fontsize = 14, fontface="bold", col = colors[colnames(X.U)]), row_names_gp = gpar(fontsize = 14, fontface="bold", col = FunCatPal[FunCat.annotation]), column_title_gp = gpar(fontsize = 18, fontface="bold"), row_title_gp = gpar(fontsize = 18, fontface="bold"), row_names_max_width = unit(150, "cm"), column_names_max_height = unit(150, "cm"))+
Heatmap(X.D, rect_gp = gpar(col = "black"), name = "Down", cluster_rows = F, cluster_columns = F, col = blueCol_fun, row_names_side = "left", column_title = "Down-regulated",  column_names_gp = gpar(fontsize = 14, fontface="bold", col = colors[colnames(X.D)]), row_names_gp = gpar(fontsize = 14, fontface="bold", col = FunCatPal[FunCat.annotation]), column_title_gp = gpar(fontsize = 18, fontface="bold"), row_title_gp = gpar(fontsize = 18, fontface="bold"), right_annotation = ha_rows, row_names_max_width = unit(150, "cm"), column_names_max_height = unit(150, "cm"))
dev.off()

```

# Generate supplementary plots

## Volcano plot
### Excitatory neurons
```{r}
require(EnhancedVolcano)

idx = grep("^Ex", colnames(DE.new$DE.sc))
Grobs = vector("list", length(idx))
for(i in 1:length(idx)) {
  k = idx[[i]]
  df = data.frame("log2FoldChange" = DE.new$logFC[, k], "pvalue" = 10^(-DE.new$logPvals[, k]))
  rownames(df) = rownames(DE.new$DE.sc)
  df = df[df$log2FoldChange != 0, ]
  
  keyvals <- rep('#cccccc', nrow(df))
  names(keyvals) <- rep('None', nrow(df))
  
  keyvals[which( (df$log2FoldChange > 0.1) & (df$pvalue < 0.05) )] <- '#ca0020'
  names(keyvals)[which( (df$log2FoldChange > 0.1) & (df$pvalue < 0.05) )] <- rep('Up', sum(keyvals == '#ca0020'))
  
  keyvals[which( (df$log2FoldChange < -0.1) & (df$pvalue < 0.05) )] <- '#0571b0'
  names(keyvals)[which( (df$log2FoldChange < -0.1) & (df$pvalue < 0.05) )] <- rep('Down', sum(keyvals == '#0571b0'))
  
  
 
 Grobs[[i]] = EnhancedVolcano(df,
  lab = rownames(df),
  x = 'log2FoldChange',
  y = 'pvalue', pCutoff = 0.05, FCcutoff = 0.1, xlim = c(-0.6, 0.6), ylim = c(0, 5), title = "Excitatory neurons", subtitle = colnames(DE.new$logFC)[[k]], colCustom = keyvals, labCol = 'black',
    labFace = 'bold', caption = "")
}  


pdf(file.path(figures.path, "Supp", "Volcano_Ex.pdf"), width = 8*4, height = 8*3)
# png(file.path(figures.path, "Supp", "Volcano_Ex.png"), width = 2400, height = 1800, res = 150)
gridExtra::grid.arrange( grobs = Grobs, nrow = 3)
dev.off()

```

### Inhibitory neurons
```{r}
require(EnhancedVolcano)


idx = grep("^In", colnames(DE.new$DE.sc))
Grobs = vector("list", length(idx))
for(i in 1:length(idx)) {
  k = idx[[i]]
  df = data.frame("log2FoldChange" = DE.new$logFC[, k], "pvalue" = 10^(-DE.new$logPvals[, k]))
  rownames(df) = rownames(DE.new$DE.sc)
  df = df[df$log2FoldChange != 0, ]
  
  keyvals <- rep('#cccccc', nrow(df))
  names(keyvals) <- rep('None', nrow(df))
  
  keyvals[which( (df$log2FoldChange > 0.1) & (df$pvalue < 0.05) )] <- '#ca0020'
  names(keyvals)[which( (df$log2FoldChange > 0.1) & (df$pvalue < 0.05) )] <- rep('Up', sum(keyvals == '#ca0020'))
  
  keyvals[which( (df$log2FoldChange < -0.1) & (df$pvalue < 0.05) )] <- '#0571b0'
  names(keyvals)[which( (df$log2FoldChange < -0.1) & (df$pvalue < 0.05) )] <- rep('Down', sum(keyvals == '#0571b0'))
  
  
 
 Grobs[[i]] = EnhancedVolcano(df,
  lab = rownames(df),
  x = 'log2FoldChange',
  y = 'pvalue', pCutoff = 0.05, FCcutoff = 0.1, xlim = c(-0.6, 0.6), ylim = c(0, 5), title = "Inhibitory neurons", subtitle = colnames(DE.new$logFC)[[k]], colCustom = keyvals, labCol = 'black',
    labFace = 'bold', caption = "")
}  



pdf(file.path(figures.path, "Supp", "Volcano_In.pdf"), width = 8*4, height = 8*2)
# png(file.path(figures.path, "Supp", "Volcano_Ex.png"), width = 2400, height = 1800, res = 150)
gridExtra::grid.arrange( grobs = Grobs, nrow = 2)
dev.off()


```

### Glial cell types
```{r}
require(EnhancedVolcano)


idx = which(!grepl("^In|^Ex", colnames(DE.new$DE.sc)))
Grobs = vector("list", length(idx))
for(i in 1:length(idx)) {
  k = idx[[i]]
  df = data.frame("log2FoldChange" = DE.new$logFC[, k], "pvalue" = 10^(-DE.new$logPvals[, k]))
  rownames(df) = rownames(DE.new$DE.sc)
  df = df[df$log2FoldChange != 0, ]
  
  keyvals <- rep('#cccccc', nrow(df))
  names(keyvals) <- rep('None', nrow(df))
  
  keyvals[which( (df$log2FoldChange > 0.1) & (df$pvalue < 0.05) )] <- '#ca0020'
  names(keyvals)[which( (df$log2FoldChange > 0.1) & (df$pvalue < 0.05) )] <- rep('Up', sum(keyvals == '#ca0020'))
  
  keyvals[which( (df$log2FoldChange < -0.1) & (df$pvalue < 0.05) )] <- '#0571b0'
  names(keyvals)[which( (df$log2FoldChange < -0.1) & (df$pvalue < 0.05) )] <- rep('Down', sum(keyvals == '#0571b0'))
  
  
 
 Grobs[[i]] = EnhancedVolcano(df,
  lab = rownames(df),
  x = 'log2FoldChange',
  y = 'pvalue', pCutoff = 0.05, FCcutoff = 0.1, xlim = c(-0.6, 0.6), ylim = c(0, 5), title = "Non-neuronal", subtitle = colnames(DE.new$logFC)[[k]], colCustom = keyvals, labCol = 'black',
    labFace = 'bold', caption = "")
}  

pdf(file.path(figures.path, "Supp", "Volcano_Glial.pdf"), width = 8*3, height = 8*1)
# png(file.path(figures.path, "Supp", "Volcano_Ex.png"), width = 2400, height = 1800, res = 150)
gridExtra::grid.arrange( grobs = Grobs, nrow = 1)
dev.off()


```

# Functional enrichment (v2.0)
## Perform functional enrichment using gProfiler
```{r}
GS = c(Up.genes, Down.genes)
names(GS) = c(paste("Up_", names(Up.genes), sep = ""),paste("Down_", names(Down.genes), sep = ""))

DE.gp = gprofiler2::gost(GS, exclude_iea = TRUE, multi_query = T, source = c("GO:BP"))
readr::write_rds(DE.gp, file.path(dataset.path, "DE_gProfiler_enrichment.rds"))

tbl = DE.gp$result[ -11]
Up.DE.tbls = lapply(1:20, function(k) {
  print(k)
  parts = apply(tbl, 2, function(x) {
    if(min(sapply(x, length)) > 1) {
      z = lapply(x, function(y) y[[k]])
    } else {
      return(x)
    }
  })
  
  sub.tbl = as.data.frame(do.call(cbind, parts))
  sub.tbl = sub.tbl[order(as.numeric(sub.tbl$p_values)), ]
  sub.tbl = sub.tbl[sub.tbl$significant == T, ]
  
  return(sub.tbl)
})
names(Up.DE.tbls) = names(Up.genes)
Up.DE.tbls = Up.DE.tbls[sapply(Up.DE.tbls, nrow) > 0]

Down.DE.tbls = lapply(21:40, function(k) {
  print(k)
  parts = apply(tbl, 2, function(x) {
    if(min(sapply(x, length)) > 1) {
      z = lapply(x, function(y) y[[k]])
    } else {
      return(x)
    }
  })
  
  sub.tbl = as.data.frame(do.call(cbind, parts))
  sub.tbl = sub.tbl[order(as.numeric(sub.tbl$p_values)), ]
  sub.tbl = sub.tbl[sub.tbl$significant == T, ]
  
  return(sub.tbl)
})
names(Down.DE.tbls) = names(Down.genes)
Down.DE.tbls = Down.DE.tbls[sapply(Down.DE.tbls, nrow) > 0]


```


## Export gProfiler results as excel tables
```{r}
Up.wb <- createWorkbook()
for(i in 1:length(Up.DE.tbls)) {
  res = Up.DE.tbls[[i]]
  n = names(Up.DE.tbls)[[i]]
  
  addWorksheet(wb=Up.wb, sheetName = n)
  writeData(Up.wb, sheet = n, res) 

}
saveWorkbook(Up.wb, file.path(tables.path, "Up_DE_gProfiler.xlsx"), overwrite = TRUE)


Down.wb <- createWorkbook()
for(i in 1:length(Down.DE.tbls)) {
  res = Down.DE.tbls[[i]]
  n = names(Down.DE.tbls)[[i]]
  
  addWorksheet(wb=Down.wb, sheetName = n)
  writeData(Down.wb, sheet = n, res) 

}
saveWorkbook(Down.wb, file.path(tables.path, "Down_DE_gProfiler.xlsx"), overwrite = TRUE)

```

## Identify meta-terms / GO clusters
Neurodevelopment
Regulation of AMPA receptor activity/regulation of NMDA receptor activity
Synaptic signaling
Oxidative phosphorylation
Synapse organization
Chaperone-mediated autophagy
Response to axon injury/response to heat
Regulation of postsynaptic membrane potential
Learning/memory
Cognition
[Regulation of catalytic activity]
Negative regulation of apoptosis


```{r}
min.GO.size = 3

DE.gp.enrichment = -log10(t(do.call(cbind, DE.gp$result$p_values)))
rownames(DE.gp.enrichment) = DE.gp$result$term_id

library(simplifyEnrichment)
GOSemSim = GO_similarity(DE.gp$result$term_id, ont = "BP")


set.seed(0)
pdf(file.path(figures.path, "DE_gp_SemSim_min3.pdf"), width = 10)
GOdf = simplifyGO(GOSemSim, min_term = min.GO.size, order_by_size = T)
dev.off()

cl.size = sapply(split(1:nrow(GOdf), GOdf$cluster), length)
GOdf$cluster = factor(GOdf$cluster, order(cl.size, decreasing = T))

GOdf2 = GOdf[order(GOdf$cluster), ]
GOdf2 = GOdf2[GOdf2$cluster %in% which(cl.size >= min.GO.size), ]

GOdf2$GO_module = match(GOdf2$cluster, unique(GOdf2$cluster))
GOdf2$max_enrichment = apply(DE.gp.enrichment[GOdf2$id, ], 1, max)

write.table(GOdf2, file.path(tables.path, "Combined_DE_BP_clusters_min3.tsv"), sep = "\t", row.names = F, col.names = T, quote = F)

```



## Combine enrichment of terms within each GO cluster
```{r}
IDX = split(GOdf2$id, GOdf2$GO_module)
combined_enrichment = do.call(rbind, lapply(IDX, function(idx) combine.logPvals(DE.gp.enrichment[idx, ])))
combined_enrichment[combined_enrichment < 0] = 0
combined_enrichment.corrected = -log10(matrix(p.adjust(10^(-combined_enrichment)), ncol = 40))
# combined_enrichment.corrected = -log10(apply(10^(-combined_enrichment), 2, function(p) p.adjust(p, "fdr")))

combined_enrichment.corrected[combined_enrichment.corrected < -log10(0.05)] = 0

combined_enrichment.corrected.up = combined_enrichment.corrected[, 1:20]
colnames(combined_enrichment.corrected.up) = names(Up.genes)
combined_enrichment.corrected.up = combined_enrichment.corrected.up[, fast_column_sums(combined_enrichment.corrected.up) > 0]

combined_enrichment.corrected.down = combined_enrichment.corrected[, 21:40]
colnames(combined_enrichment.corrected.down) = names(Down.genes)
combined_enrichment.corrected.down = combined_enrichment.corrected.down[, fast_column_sums(combined_enrichment.corrected.down) > 0]


ha_rows = rowAnnotation(df = list("Class" = FunCat.annotation), col = list("Class" = FunCatPal), annotation_legend_param = list("Class"=list(title_gp = gpar(fontsize = 0), labels_gp = gpar(fontsize = 10))))


X.U = (combined_enrichment.corrected.up)
redCol_fun = circlize::colorRamp2(c(0, exp(quantile(log(X.U)[X.U > -log10(0.05)], seq(0.05, 0.95, length.out = 12)))), c("#ffffff", pals::brewer.reds(12)))

X.D = (combined_enrichment.corrected.down)
blueCol_fun = circlize::colorRamp2(c(0, exp(quantile(log(X.D)[X.D > -log10(0.05)], seq(0.05, 0.95, length.out = 12)))), c("#ffffff", pals::brewer.blues(12)))

pdf(file.path(figures.path, "DE_gProfiler_Simplified.pdf"), width = 5, height = 6)
par(mar=c(0,150,0,0))
Heatmap(X.U, rect_gp = gpar(col = "black"), name = "Up", column_title = "Up", cluster_rows = F, cluster_columns = F, col = redCol_fun, row_names_side = "left",  column_names_gp = gpar(fontsize = 14, fontface="bold", col = colors[colnames(X.U)]), row_names_gp = gpar(fontsize = 14, fontface="bold", col = FunCatPal[FunCat.annotation]), column_title_gp = gpar(fontsize = 18, fontface="bold"), row_title_gp = gpar(fontsize = 18, fontface="bold"), row_names_max_width = unit(150, "cm"), column_names_max_height = unit(150, "cm"))+
Heatmap(X.D, rect_gp = gpar(col = "black"), name = "Down", cluster_rows = F, cluster_columns = F, col = blueCol_fun, row_names_side = "left", column_title = "Down",  column_names_gp = gpar(fontsize = 14, fontface="bold", col = colors[colnames(X.D)]), row_names_gp = gpar(fontsize = 14, fontface="bold", col = FunCatPal[FunCat.annotation]), column_title_gp = gpar(fontsize = 18, fontface="bold"), row_title_gp = gpar(fontsize = 18, fontface="bold"), row_names_max_width = unit(150, "cm"), column_names_max_height = unit(150, "cm"))
dev.off()


```


