---
title: "H-MAGMA analysis"
output: html_notebook
---

# Setup
```{r include=FALSE}
library(org.Hs.eg.db)
require(ACTIONet)
require(stringr)
require(ComplexHeatmap)


results.path = "~/results"
input.path = "~/results/input"
dataset.path = "~/results/datasets"
tables.path = "~/results/tables"
figures.path = "~/results/figures"
MAGMA.path = "~/magma"


```

## Enrichment function
```{r}
assess.genesets <-function (arch.gs, terms.gs, N, min.pval = 1e-100, correct = "none"){
    shared = t(sapply(terms.gs, function(gs1) {
        sapply(arch.gs, function(gs2) {
            nn = intersect(gs1, gs2)
        })
    }))
    colnames(shared) = names(arch.gs)
    GS.sizes = sapply(terms.gs, length)
    logPvals.out = sapply(1:ncol(shared), function(i) {
        gg = shared[, i]
        x = as.numeric(sapply(gg, length))
        n.sample = length(arch.gs[[i]])
        n.success = as.numeric(GS.sizes)
        v = rep(1, length(x))
        min.overlap = n.success * n.sample/N
        idx = which(x >= min.overlap)
        if (length(idx) == 0) 
            return(v)
        v[idx] = (phyper(x[idx]-1, n.sample, N-n.sample, n.success[idx], lower.tail = F))
        # v[idx] = HGT_tail(population.size = N, success.count = n.success[idx], 
        #     sample.size = n.sample, observed.success = x[idx])
        return(v)
    })
    if(correct == "global") {
      logPvals.out = matrix(p.adjust(logPvals.out, "fdr"), nrow = nrow(logPvals.out))
    } else if(correct == "local") {
      logPvals.out = apply(logPvals.out, 2, function(x) p.adjust(x, "fdr"))
    }
    rownames(logPvals.out) = names(terms.gs)
    colnames(logPvals.out) = names(arch.gs)
    return(-log10(Matrix::t(logPvals.out)))
}
```

# Load primary datasets
```{r, eval = T}
ACTIONet_summary = readr::read_rds(file.path(dataset.path, "ACTIONet_summary_filtered_individuals.rds"))
pb.logcounts = readr::read_rds(file.path(dataset.path, "PB_mean_logcounts_final.RDS"))

color.df = readRDS(file.path(dataset.path, "celltype_colors.rds"))
colors = color.df$color
names(colors) = color.df$celltype

```


# Load DE results
```{r, eval = T}
resDE = readr::read_rds(file.path(dataset.path, "Cohort_specific_DE_results.rds"))
filtered.tables = readr::read_rds(file.path(dataset.path, "Cohort_specific_DE_results_filtered.rds"))
combined.analysis.tables = readr::read_rds(file.path(dataset.path, "meta_analysis_results.rds"))

DE.new = readRDS(file.path(dataset.path, "DE_genes_pseudobulk.rds"))
Up.genes = DE.new$Up.genes
Down.genes = DE.new$Down.genes
DE.sc = DE.new$DE.sc
ordered.celltypes = rownames(X)[order(apply(X, 1, sum), decreasing = T)]

```



# Convert DE genes to ENSG and export
```{r}
require(org.Hs.eg.db)

suppressWarnings(ids <- AnnotationDbi::mapIds(org.Hs.eg.db, keys = row.names(DE.new$DE.sc), keytype = "SYMBOL", column = "ENSEMBL", multiVals = "first"))
ids[is.na(ids)] = ""

Up.genes.ENSG = sapply(Up.genes, function(gs) {
  setdiff(sort(unique(ids[match(gs, rownames(pb.logcounts))])), "")
})
Up.genes.ENSG.df = reshape2::melt(Up.genes.ENSG)
Up.genes.ENSG.df = Up.genes.ENSG.df[, c(2, 1)]
Up.genes.ENSG.df$L1 = factor(Up.genes.ENSG.df$L1, names(Up.genes))
Up.genes.ENSG.df = Up.genes.ENSG.df[order(Up.genes.ENSG.df$L1), ]
write.table(Up.genes.ENSG.df, file.path(MAGMA.path, "genesets", "Up_genes_ENSG.tsv"), sep = "\t", col.names = F, row.names = F, quote = F)

Down.genes.ENSG = sapply(Down.genes, function(gs) {
  setdiff(sort(unique(ids[match(gs, rownames(pb.logcounts))])), "")
})
Down.genes.ENSG.df = reshape2::melt(Down.genes.ENSG)
Down.genes.ENSG.df = Down.genes.ENSG.df[, c(2, 1)]
Down.genes.ENSG.df$L1 = factor(Down.genes.ENSG.df$L1, names(Down.genes))
Down.genes.ENSG.df = Down.genes.ENSG.df[order(Down.genes.ENSG.df$L1), ]

write.table(Down.genes.ENSG.df, file.path(MAGMA.path, "genesets", "Down_genes_ENSG.tsv"), sep = "\t", col.names = F, row.names = F, quote = F)


UpAndDown.genes.ENSG.df = rbind(Up.genes.ENSG.df, Down.genes.ENSG.df)
UpAndDown.genes.ENSG.df$L1 = factor(UpAndDown.genes.ENSG.df$L1, names(Up.genes))
UpAndDown.genes.ENSG.df = UpAndDown.genes.ENSG.df[order(UpAndDown.genes.ENSG.df$L1), ]

write.table(UpAndDown.genes.ENSG.df, file.path(MAGMA.path, "genesets", "UpAndDown_genes_ENSG.tsv"), sep = "\t", col.names = F, row.names = F, quote = F)


```





# Run H-MAGMA
We run HMAGMA using  **for f in `ls hmagma/`; do (./magma --gene-results ./hmagma/$f/$f.genes.raw --set-annot genesets/UpAndDown_genes_ENSG.tsv col=2,1 --out $f\_UpAndDown &); done`**, from inside the magma folder, where hmagma are prestored the gene models scores.

# Load H-MAGMA results

```{r}
dl = list.dirs(file.path(MAGMA.path, "psych_arch/"), full.names = F, recursive = F)

HMAGMA.Pvals = vector("list", 1)
names(HMAGMA.Pvals) = c("UpAndDown")
for (d in names(HMAGMA.Pvals)) {
  HMAGMA.Pvals[[d]] = matrix(1, nrow = 20, ncol = length(dl))
  colnames(HMAGMA.Pvals[[d]]) = dl
  rownames(HMAGMA.Pvals[[d]]) = names(Up.genes)
  for(cond in dl) {
    print(cond)
    file.name = sprintf("%s/out/%s_%s.gsa.out", MAGMA.path, cond, d)
    lines = readLines(con <- file(file.name))
    lines = str_split(lines, "\n")[-c(1:5)]
    
    pvals = sapply(lines, function(ll) {
      parts = str_split(ll, " ")
      as.numeric(parts[[1]][length(parts[[1]])])
    })
    
    names(pvals) = sapply(lines, function(ll) {
      parts = str_split(ll, " ")
      parts[[1]][[1]]
    })

    HMAGMA.Pvals[[d]][names(pvals), cond] = pvals 
  }
}


readr::write_rds(HMAGMA.Pvals, file.path(dataset.path, "HMAGMA_results_raw.rds"))


```

# Or load preprocessed results
```{r}
HMAGMA.Pvals = readr::read_rds(file.path(dataset.path, "HMAGMA_results_raw.rds"))

```



## Export selected results
```{r}

  selected.traits = c("hmagmaAdultBrain__sz3", "hmagmaAdultBrain__bip2", "hmagmaAdultBrain__asd", "hmagmaAdultBrain__adhd", "hmagmaAdultBrain__mdd_without_23andMe", "hmagmaAdultBrain__alz2noapoe")

trait.labels = c("Schizophrenia (SZ)", "Bipolar (BP)","Autism (ASD)", "ADHD", "Depression (MDD)", "Alzheimer (AD)")

X = HMAGMA.Pvals$UpAndDown[, selected.traits]
cts = intersect(color.df$celltype, rownames(X))
X = X[cts, ]
X = matrix(p.adjust(as.numeric(X), "fdr"), nrow = length(cts))
rownames(X) = cts
# X = apply(X, 2, function(p) p.adjust(p, "fdr"))
rownames(X) = intersect(names(colors), rownames(X))
X = -log10(X)
X[X < -log10(0.05)] = 0
colnames(X) = trait.labels
rownames(X) = cts

PurPal = colorRampPalette(RColorBrewer::brewer.pal(9, "Purples"))(200)
PurPal = c(rep(PurPal[1], length(PurPal)*(sum(X < 1) / length(X))), PurPal)

require(corrplot)
pdf(file.path(figures.path, "HMAGMA_adultBrain.pdf"), width =7, height = 7)
corrplot(X, is.corr = F, method = "pie", col = PurPal, cl.lim = c(0, 4), cl.length = 5, outline = T, sig.level = 0.05, p.mat = 10^(-X), insig = "blank", tl.col = "black")
dev.off()

X = X[grep("^In|^Ex", rownames(X)), ]
require(corrplot)
pdf(file.path(figures.path, "HMAGMA_adultBrain_neuro.pdf"), width =7, height = 7)
corrplot(X, is.corr = F, method = "pie", col = PurPal, cl.lim = c(0, 4), cl.length = 5, outline = T, sig.level = 0.05, p.mat = 10^(-X), insig = "blank", tl.col = "black")
dev.off()

```

# Plot "volcanos"
## Load H-MAGMA gene p-values
```{r}
tbl = read.table("~/magma/hmagma/hmagmaAdultBrain__sz3/hmagmaAdultBrain__sz3.genes.out", header = T)
suppressWarnings(ids <- AnnotationDbi::mapIds(org.Hs.eg.db, keys = tbl$GENE, keytype = "ENSEMBL", column = "SYMBOL", multiVals = "first"))
ids[is.na(ids)] = ""

scores = rep(0, nrow(pb.logcounts))
ii = match(rownames(pb.logcounts), ids)
scores[!is.na(ii)] = -log10(tbl$P[ii[!is.na(ii)]])
names(scores) = rownames(pb.logcounts)

sort(scores, decreasing = T)[1:30]

```

```{r}
selected.cts = c("Ex-L45_LRRK1", "In-PV_Basket", "Oli")
idx = match(selected.cts, colnames(DE.new$DE.sc))

t.threshold = 3
magma.pval.threshold = 0.05

Grobs = vector("list", length(idx))
for(i in 1:length(idx)) {
  k = idx[[i]]
  df = data.frame("log2FoldChange" = DE.new$DE.sc[, k], "pvalue" = 10^(-scores))
  rownames(df) = rownames(DE.new$DE.sc)
  df = df[df$log2FoldChange != 0, ]
  
  keyvals <- rep('#cccccc', nrow(df))
  names(keyvals) <- rep('None', nrow(df))
  
  keyvals[which( (df$log2FoldChange > t.threshold) & (df$pvalue < magma.pval.threshold) )] <- '#ca0020'
  names(keyvals)[which( (df$log2FoldChange > t.threshold) & (df$pvalue < magma.pval.threshold) )] <- rep('Up', sum(keyvals == '#ca0020'))
  
  keyvals[which( (df$log2FoldChange < -t.threshold) & (df$pvalue < magma.pval.threshold) )] <- '#0571b0'
  names(keyvals)[which( (df$log2FoldChange < -t.threshold) & (df$pvalue < magma.pval.threshold) )] <- rep('Down', sum(keyvals == '#0571b0'))
  
  
 
 Grobs[[i]] = EnhancedVolcano(df,
  lab = rownames(df),
  x = 'log2FoldChange',
  y = 'pvalue', pCutoff = magma.pval.threshold, FCcutoff = t.threshold, xlim = c(-4, 4), ylim = c(0, 20), title = "H-MAGMA pvals-vs-DE tstats", subtitle = colnames(DE.new$logFC)[[k]], colCustom = keyvals, labCol = 'black',
    labFace = 'bold', caption = "")
}  

pdf(file.path(figures.path, "HMAGMA_vs_DE.pdf"), width = 8*3, height = 8*1)
gridExtra::grid.arrange( grobs = Grobs, nrow = 1)
dev.off()



```

