---
title: "Perform DE analysis"
subtitle: "Step 2: Limma-trend on PB(s) followed by meta-analysis"
output: html_notebook
---

# Setup
```{r include=FALSE}
require(ACTIONet)
require(stringr)
require(ComplexHeatmap)
require(openxlsx)

require(muscat)
require(edgeR)
require(limma)

results.path = "~/results"
input.path = "~/results/input"
dataset.path = "~/results/datasets"
tables.path = "~/results/tables"
figures.path = "~/results/figures"


dev_threshold = 1 # logFC needs to be 1 std away from 0 (in both datasets) prior to metanalysis

# Thresholds on the FDR-corrected meta-analysis results
pval_threshold = 0.05
logFC_threshold = 0.1

```




# Load datasets
```{r}
ACTIONet_summary = readr::read_rds(file.path(dataset.path, "ACTIONet_summary_filtered_individuals.rds"))
pb.logcounts = readr::read_rds(file.path(dataset.path, "PB_mean_logcounts.RDS"))

ncells = sapply(int_colData(pb.logcounts)$n_cells, as.numeric)
rownames(ncells) = names(assays(pb.logcounts))

```



# Prefilter outlier samples using % of excitatory neurons
SZ33 is removed due to having > 80% ExNeu, and samples SZ3, SZ15, SZ24, SZ29 are removed due to having less than 10% ExNeu
```{r}
cs = Matrix::colSums(ncells)
ncells.freq = 100*scale(ncells, center = F, scale = cs)
Ex.perc = (fast_column_sums(ncells.freq[grepl("^Ex", rownames(ncells.freq)) & !grepl("^Ex-NRGN", rownames(ncells.freq)), ]))

mask = (Ex.perc >= 10) & (Ex.perc <= 80) 
pb.logcounts.filtered = pb.logcounts [, mask]

```



# Performing cohort-specific DE
```{r}
pb.logcounts.filtered$SampleQuality = scale(log1p(pb.logcounts.filtered$umis))
form = ~ Phenotype + Batch + PMI + Gender + Age + Benzodiazepines + Anticonvulsants + AntipsychTyp + AntipsychAtyp + Antidepress + SampleQuality

resDE = lapply( levels(pb.logcounts.filtered$Cohort), function(chrt){

	keep.ids = colnames(pb.logcounts.filtered)[pb.logcounts.filtered$Cohort == chrt]

	pb.logcounts.filtered_sub = pb.logcounts.filtered[,keep.ids]
  sample.metadata = droplevels(data.frame(colData(pb.logcounts.filtered_sub)))
	design.mat <- model.matrix(form, data = sample.metadata)
	colnames(design.mat)[1] = c("Intercept")

	contrast.mat <- makeContrasts(contrasts = "PhenotypeSZ", levels = design.mat)

	df = pbDS(pb.logcounts.filtered_sub, method = "limma-trend", min_cells = 5, design = design.mat, contrast =  contrast.mat, filter = "both")
	
})
names(resDE) = levels(colData(pb.logcounts.filtered)$Cohort)

readr::write_rds(resDE, file.path(dataset.path, "Cohort_specific_DE_results.rds"))

```


## Export as excel tables
```{r}
for(ds in 1:length(resDE)) {
  print(names(resDE)[[ds]])
  
  Up.wb <- createWorkbook()
  for(i in 1:length(resDE[[ds]]$table$PhenotypeSZ)) {
    res = resDE[[ds]]$table$PhenotypeSZ[[i]]
    res = res[res$logFC > 0, ]
    res = cbind(data.frame(Gene = rownames(res)), res)
    res = res[order(res$t, decreasing = T), ]
  
    n = names(resDE[[ds]]$table$PhenotypeSZ)[[i]]
    
    addWorksheet(wb=Up.wb, sheetName = n)
    writeData(Up.wb, sheet = n, res) 
  }
  saveWorkbook(Up.wb, sprintf(file.path(tables.path, "DE_genes_up_%s_complete_set.xlsx"), names(resDE)[[ds]]), overwrite = TRUE)
  
  
  Down.wb <- createWorkbook()
  for(i in 1:length(resDE[[ds]]$table$PhenotypeSZ)) {
    res = resDE[[ds]]$table$PhenotypeSZ[[i]]
    res = res[res$logFC < 0, ]
    res = cbind(data.frame(Gene = rownames(res)), res)
    res = res[order(res$t, decreasing = F), ]
    
    n = names(resDE[[ds]]$table$PhenotypeSZ)[[i]] 
    
    addWorksheet(wb=Down.wb, sheetName = n)
    writeData(Down.wb, sheet = n, res) 
  
  }
  saveWorkbook(Down.wb, sprintf(file.path(tables.path, "DE_genes_down_%s_complete_set.xlsx"), names(resDE)[[ds]]), overwrite = TRUE)
}

```



# Prefiltering individual DE results before combining them
Only keep genes that have 1) consistent direction of dysregulation across both datasets, and 2) at least 1 std away from zero on the logFC scale

```{r}
common.celltypes = intersect(names(resDE$McLean$table$PhenotypeSZ), names(resDE$MtSinai$table$PhenotypeSZ))

filtered.tables = lapply(common.celltypes, function(celltype) {
  tbl1 = resDE[[1]]$table$PhenotypeSZ[[celltype]]
  tbl2 = resDE[[2]]$table$PhenotypeSZ[[celltype]]
  
  genes = intersect(tbl1$gene[dev_threshold <= abs(tbl1$logFC/sd(tbl1$logFC))], tbl2$gene[dev_threshold <= abs(tbl2$logFC / sd(tbl2$logFC))])

  tbl1 = tbl1[match(genes, tbl1$gene), ]
  tbl2 = tbl2[match(genes, tbl2$gene), ]
  
  mask = sign(tbl1$logFC)*sign(tbl2$logFC) > 0
  tbl1 = tbl1[mask, ]
  tbl2 = tbl2[mask, ]
  
  tbls = list(McClean = tbl1, MtSinai = tbl2) 
  tbls = lapply( tbls, function(tab){
    tab$se = tab$logFC / tab$t
    tab
  })
  
  return(tbls)
})
names(filtered.tables) = common.celltypes

readr::write_rds(filtered.tables, file.path(dataset.path, "Cohort_specific_DE_results_filtered.rds"))

```

## Export as excel tables
```{r}
for(ds in 1:length(filtered.tables[[1]])) {
  Up.wb <- createWorkbook()
  for(i in 1:length(filtered.tables)) {
    res = filtered.tables[[i]][[ds]]
    res = res[res$logFC > 0, ]
    res = cbind(data.frame(Gene = rownames(res)), res)
    res = res[order(res$t, decreasing = T), ]
  
    n = names(filtered.tables)[[i]]
    
    addWorksheet(wb=Up.wb, sheetName = n)
    writeData(Up.wb, sheet = n, res) 
  
  }
  saveWorkbook(Up.wb, sprintf(file.path(tables.path, "DE_genes_up_%s_filtered.xlsx"), names(filtered.tables[[i]])[[ds]]), overwrite = TRUE)
  
  
  Down.wb <- createWorkbook()
  for(i in 1:length(filtered.tables)) {
    res = filtered.tables[[i]][[ds]]
    res = res[res$logFC < 0, ]
    res = cbind(data.frame(Gene = rownames(res)), res)
    res = res[order(res$t, decreasing = F), ]
    
    n = names(filtered.tables)[[i]]
    
    addWorksheet(wb=Down.wb, sheetName = n)
    writeData(Down.wb, sheet = n, res) 
  
  }
  saveWorkbook(Down.wb, sprintf(file.path(tables.path, "DE_genes_down_%s_filtered.xlsx"), names(filtered.tables[[i]])[[ds]]), overwrite = TRUE)
}

```

# Perform meta-analysis via Linear (Mixed-Effects) Models (RMA)
```{r}
combined.analysis.tables = lapply(names(filtered.tables), function(celltype) {
  print(celltype)
  tbls = filtered.tables[[celltype]]
  
  gene.tbls = lapply(1:nrow(tbls[[1]]), function(i) {
    dfs = lapply(1:length(tbls), function(k) tbls[[k]][i, ])
    df = do.call("rbind", dfs)
  })
  names(gene.tbls) = tbls[[1]]$gene
    
  combined.analysis.tbl = do.call(rbind, lapply(names(gene.tbls), function(gene){
    x = suppressWarnings(metafor::rma(yi=logFC, sei=se, data = gene.tbls[[gene]], method="FE"))
    combined.tbl = data.frame( gene = gene, 
        logFC     = x$beta,
        se        = x$se,
        tstat = x$zval,
        P.Value   = x$pval)
    return(combined.tbl)
  }))
  rownames(combined.analysis.tbl) = names(gene.tbls)
  
  combined.analysis.tbl = combined.analysis.tbl[order(combined.analysis.tbl$P.Value), ]
  
  return(combined.analysis.tbl)
})
names(combined.analysis.tables) = names(filtered.tables)

DF = do.call(rbind, combined.analysis.tables)
DF$adj.P.Val = p.adjust(DF$P.Value, "fdr")
ff = factor(unlist(lapply(names(combined.analysis.tables), function(celltype) rep(celltype, nrow(combined.analysis.tables[[celltype]])))), names(filtered.tables))
combined.analysis.tables = split(DF, ff)

readr::write_rds(combined.analysis.tables, file.path(dataset.path, "meta_analysis_results.rds"))

```


# Export final tables
```{r}
  Up.wb <- createWorkbook()
  for(i in 1:length(combined.analysis.tables)) {
    res = combined.analysis.tables[[i]]
    res = res[(res$logFC > logFC_threshold) & (res$P.Value <= pval_threshold), ]
    res = res[order(res$t, decreasing = T), ]
  
    n = names(combined.analysis.tables)[[i]] 
    
    addWorksheet(wb=Up.wb, sheetName = n)
    writeData(Up.wb, sheet = n, res) 
  
  }
  saveWorkbook(Up.wb, file.path(tables.path, "DE_genes_up_combined.xlsx"), overwrite = TRUE)
  
  
  Down.wb <- createWorkbook()
  for(i in 1:length(combined.analysis.tables)) {
    res = combined.analysis.tables[[i]]
    res = res[(res$logFC < -logFC_threshold) & (res$P.Value <= pval_threshold), ]
    res = res[order(res$t, decreasing = F), ]
    
    n = names(combined.analysis.tables)[[i]] 
    
    addWorksheet(wb=Down.wb, sheetName = n)
    writeData(Down.wb, sheet = n, res) 
  
  }
  saveWorkbook(Down.wb, file.path(tables.path, "DE_genes_down_combined.xlsx"), overwrite = TRUE)
  
```


# Summarize and simplify DE results
```{r}
DE.sc = matrix(0, nrow(pb.logcounts), length(combined.analysis.tables))
tstats = matrix(0, nrow(pb.logcounts), length(combined.analysis.tables))
logFC = matrix(0, nrow(pb.logcounts), length(combined.analysis.tables))
logPvals = matrix(0, nrow(pb.logcounts), length(combined.analysis.tables))
rownames(DE.sc) = rownames(tstats) = rownames(logFC) = rownames(logPvals) = rownames(pb.logcounts)
colnames(DE.sc) = colnames(tstats) = colnames(logFC) = colnames(logPvals) = names(combined.analysis.tables)

limma_trend_mean.scores = matrix(0, nrow(pb.logcounts), length(combined.analysis.tables))
Up.genes = vector("list", length(combined.analysis.tables))
Down.genes = vector("list", length(combined.analysis.tables))
rownames(limma_trend_mean.scores) = rownames(pb.logcounts)
names(Up.genes) = names(Down.genes) = colnames(limma_trend_mean.scores) = names(combined.analysis.tables)
for(i in 1:length(combined.analysis.tables)) {
	print(i)
	
	tbl = combined.analysis.tables[[i]]

	tstats[tbl$gene, names(combined.analysis.tables)[[i]]] = tbl$tstat
	logFC[tbl$gene, names(combined.analysis.tables)[[i]]] = tbl$logFC
	logPvals[tbl$gene, names(combined.analysis.tables)[[i]]] = -log10(tbl$adj.P.Val)
	
	x = tbl$tstat
	x[abs(tbl$logFC) < logFC_threshold] = 0
	DE.sc[tbl$gene, names(combined.analysis.tables)[[i]]] = x
	
}
limma_trend_mean.scores[is.na(limma_trend_mean.scores)] = 0


Up.genes = lapply(combined.analysis.tables, function(combined.analysis.tbl) {
  combined.analysis.tbl$gene[(combined.analysis.tbl$logFC > logFC_threshold) & (combined.analysis.tbl$adj.P.Val < pval_threshold)]
})
Down.genes = lapply(combined.analysis.tables, function(combined.analysis.tbl) {
  combined.analysis.tbl$gene[(combined.analysis.tbl$logFC < -logFC_threshold) & (combined.analysis.tbl$adj.P.Val < pval_threshold)]
})

DE.new = list(DE.sc = DE.sc, tstats = tstats, logFC = logFC, logPvals = logPvals, Up.genes = Up.genes, Down.genes = Down.genes)

saveRDS(DE.new, file.path(dataset.path, "DE_genes_pseudobulk.rds"))



```


