---
title: "Annotate and verify cell type annotations"
output: html_notebook
---

# Setup
```{r include=FALSE}
require(ACTIONet)
require(stringr)
require(ComplexHeatmap)

dataset.path = "~/results/datasets/"
results.path = "~/results"
figures.path = "~/results/figures"


```

# Load summary data
```{r}
pb.logcounts = readr::read_rds(file.path(dataset.path, "PB_sce_final.rds"))

ACTIONet_summary = readr::read_rds(file.path(dataset.path, "ACTIONet_summary_filtered_individuals.rds"))

celltype.gene.spec = readr::read_rds(file.path(dataset.path, "celltype_gene_specificity.rds"))
archetype.gene.spec = readr::read_rds(file.path(dataset.path, "archetype_gene_specificity.rds"))

color.df = readRDS(file.path(dataset.path, "celltype_colors.rds"))




colors = color.df$color
names(colors) = color.df$celltype

```


# Garbage!
```{r}
ace = readr::read_rds("~/results/ACTIONet_reunified.rds")

ace = rerun.archetype.aggregation(ace)

meta = readRDS("~/PFC_v3/SZ_cell_meta.rds")


colors = readRDS("~/results/celltype_colors.rds")

ll = c("Ex-NRGN", "Ex-L2", "Ex-L23", "Ex-L3", "Ex-L4_MYLK", "Ex-L45_MET", "Ex-L45_LRRK1", "Ex-L5b_VEN", "Ex-L5b_HTR2C", "Ex-L56", "Ex-L56_CC_NTNG2", "Ex-L6_CC_SEMA3A", "Ex-L6b_SEMA3D", "Ex-L6b_SEMA3E", "In-Rosehip_CHST9", "In-Rosehip_TRPC3", "In-Reelin", "In-VIP", "In-PV_Chandelier", "In-PV_Basket", "In-SST", "Oli", "OPC", "Ast", "Mic", "Endo", "Pericytes")


ll.old = c("Ex-NRGN", "Ex-L2", "Ex-L23", "Ex-L3", "Ex-L4_MYLK", "Ex-L45", "Ex-L45_LRRK1", "Ex-L5b", "Ex-L5b_HTR2C", "Ex-L5", "Ex-CC_THEMIS_NTNG2", "Ex-CC_THEMIS", "Ex-L6_SEMA3D", "Ex-L6", "In-Rosehip_CHST9", "In-Rosehip_TRPC3", "In-Reelin", "In-VIP", "In-PV_Chandelier", "In-PV_Basket", "In-SST", "Oli", "OPC", "Ast", "Mic", "Endothelial cells", "Pericytes")

names(colors)[[23]] = "Endothelial cells"

cl = as.character(colors[ll.old])
# cl[1:14] = pals::brewer.greens(16*3)[seq(3*3, 16*3, 3)]
cl[1:14] = rev(pals::kovesi.linear_green_5_95_c69(17)[c(1, 4, 5:17)])
cl[15:21] = pals::brewer.reds(9)[2:8]
new.colors = cl
names(new.colors) = ll

Labels = factor(ll[match(ace$Labels.final, ll.old)], ll)





# color.df = readRDS("~/results/datasets/celltype_colors.rds")

# ace$Labels = Labels

# df = data.frame(celltype = ll, old.celltype = ll.old, color = cl)
# readr::write_rds(df, file = "~/results/datasets/celltype_colors.rds")

X = data.frame(colData(ace))[, c(1:7, 14, 20)]

ACTIONet_summary = list(metadata = X, ACTIONet2D = ace$ACTIONet2D, ACTIONet3D = ace$ACTIONet3D, H_unified = colMaps(ace)$H_unified, unified_feature_specificity = ace$unified_feature_specificity)
ACTIONet_summary$metadata$umi_count = meta$umis
ACTIONet_summary$metadata$gene_counts = meta$genes
ACTIONet_summary$metadata$mito_perc = meta$mito_perc
ACTIONet_summary$metadata$Labels = Labels
ACTIONet_summary$metadata$Batch = pb.logcounts$Batch[ACTIONet_summary$metadata$Individual]
ACTIONet_summary$metadata$Gender = pb.logcounts$Gender[ACTIONet_summary$metadata$Individual]
ACTIONet_summary$metadata$Age = pb.logcounts$Age[ACTIONet_summary$metadata$Individual]
ACTIONet_summary$metadata$PMI = pb.logcounts$PMI[ACTIONet_summary$metadata$Individual]


ACTIONet_summary$metadata$dataset[ACTIONet_summary$metadata$dataset == "DS1"] = "McLean"
ACTIONet_summary$metadata$dataset[ACTIONet_summary$metadata$dataset == "DS2"] = "MtSinai"
ACTIONet_summary$metadata$assigned_archetype = ace$assigned_archetype

ACTIONet_summary$metadata$Individual = pb.logcounts$ID[match(ACTIONet_summary$metadata$Individual, pb.logcounts$Internal_ID)]

readr::write_rds(ACTIONet_summary, "~/results/datasets/ACTIONet_summary.rds")

na.mask = is.na(ACTIONet_summary$metadata$Individual)
ACTIONet_summary$metadata = ACTIONet_summary$metadata[!na.mask, ]
ACTIONet_summary$ACTIONet2D = ACTIONet_summary$ACTIONet2D[!na.mask, ]
ACTIONet_summary$ACTIONet3D = ACTIONet_summary$ACTIONet3D[!na.mask, ]
ACTIONet_summary$H_unified = ACTIONet_summary$H_unified[!na.mask, ]

readr::write_rds(ACTIONet_summary, "~/results/datasets/ACTIONet_summary_filtered_individuals.rds")



```

```{r}
ACTIONet_summary = readr::read_rds("~/results/datasets/ACTIONet_summary.rds")

```

# Plot main ACTIONet
## Celltypes
```{r}
# annotated
gg1 = my.plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$Labels, palette = new.colors, text_size = 2, use_repel = T)

png(file.path(figures.path, "ACTIONet_annotated.png"), width = 1600, height = 1200, res = 150)
plot(gg1)
dev.off()

gg2 = my.plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$Labels, palette = new.colors, text_size = 2, use_repel = T, add_text_labels = F)

png(file.path(figures.path, "ACTIONet_annotated_no_labels.png"), width = 1600, height = 1200, res = 150)
plot(gg2)
dev.off()

```


## Supps
### # Phenotype
```{r}
gg2 = my.plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$Phenotype, palette = c("#cccccc", "#888888"), text_size = 2, use_repel = T, add_text_labels = F, point_size = 0.5)

png(file.path(figures.path, "ACTIONet_phenotype_annotated_no_labels.png"), width = 1600, height = 1200, res = 150)
plot(gg2)
dev.off()
```

### Batch
```{r}
gg2 = my.plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$Batch, text_size = 2, use_repel = T, add_text_labels = F, point_size = 0.5)

png(file.path(figures.path, "ACTIONet_batch_annotated_no_labels.png"), width = 1600, height = 1200, res = 150)
plot(gg2)
dev.off()
```

### Gender
```{r}
gg2 = my.plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$Gender, text_size = 2, use_repel = T, add_text_labels = F, point_size = 0.5, palette = c("pink", "#91bfdb"))

png(file.path(figures.path, "ACTIONet_gender_annotated_no_labels.png"), width = 1600, height = 1200, res = 150)
plot(gg2)
dev.off()
```

## Archetype
```{r}
gg2 = my.plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$assigned_archetype, text_size = 2, use_repel = T, add_text_labels = F, point_size = 0.5, palette = as.character(pals::polychrome(31)))

png(file.path(figures.path, "ACTIONet_archetypes_annotated_no_labels.png"), width = 1600, height = 1200, res = 150)
plot(gg2)
dev.off()
```
## Archetype
```{r}
gg2 = plot.ACTIONet(ACTIONet_summary$ACTIONet2D, ACTIONet_summary$metadata$dataset, text_size = 2, use_repel = T, add_text_labels = F, point_size = 0.5, palette = c("#f1a340", "#998ec3"))

png(file.path(figures.path, "ACTIONet_datasets_annotated_no_labels.png"), width = 1600, height = 1200, res = 150)
plot(gg2)
dev.off()

```

```{r}
ct.vs.ds = table(interaction(ACTIONet_summary$metadata$Labels, ACTIONet_summary$metadata$dataset))
cc = matrix(as.numeric(ct.vs.ds), ncol = 2)
cs = fast_column_sums(cc)
cc = 100*scale(cc, center = F, scale = cs)
rownames(cc) = levels(ACTIONet_summary$metadata$Labels)
Heatmap(log2(cc[, 1]/cc[, 2]))

```


## Plot ACTIONet plots per dataset
### McLean dataset
```{r}
mask = ACTIONet_summary$metadata$dataset == "McLean"
DS1.coors = ACTIONet_summary$ACTIONet2D[mask, ]
DS1.labels = ACTIONet_summary$metadata$Labels[mask]

gg3 = my.plot.ACTIONet(DS1.coors, DS1.labels, palette = new.colors, text_size = 2, use_repel = T)

png(file.path(figures.path, "ACTIONet_annotated_DS1.png"), width = 1600, height = 1200, res = 150)
plot(gg3)
dev.off()

gg4 = my.plot.ACTIONet(DS1.coors, DS1.labels, palette = new.colors, text_size = 2, use_repel = T, add_text_labels = F)

png(file.path(figures.path, "ACTIONet_annotated_no_labels_DS1.png"), width = 1600, height = 1200, res = 150)
plot(gg4)
dev.off()

```

### MtSinai dataset
```{r}
mask = ACTIONet_summary$metadata$dataset == "MtSinai"
DS2.coors = ACTIONet_summary$ACTIONet2D[mask, ]
DS2.labels = ACTIONet_summary$metadata$Labels[mask]

gg5 = my.plot.ACTIONet(DS1.coors, DS1.labels, palette = new.colors, text_size = 2, use_repel = T)

png(file.path(figures.path, "ACTIONet_annotated_DS2.png"), width = 1600, height = 1200, res = 150)
plot(gg5)
dev.off()

gg6 = my.plot.ACTIONet(DS1.coors, DS1.labels, palette = new.colors, text_size = 2, use_repel = T, add_text_labels = F)

png(file.path(figures.path, "ACTIONet_annotated_no_labels_DS2.png"), width = 1600, height = 1200, res = 150)
plot(gg6)
dev.off()

```


# Plot cell type fraction stats
```{r}
require(ggpubr)

ncells = sapply(int_colData(pb.logcounts)$n_cells, as.numeric)
rownames(ncells) = names(assays(pb.logcounts))

cs = Matrix::colSums(ncells)
ncells.freq = 100*scale(ncells, center = F, scale = cs)
# Ex.perc = (fast_column_sums(ncells.freq[grepl("^Ex", rownames(ncells.freq)) & !grepl("^Ex-NRGN", rownames(ncells.freq)), ]))
Ex.perc = (fast_column_sums(ncells.freq[grepl("^Ex", rownames(ncells.freq)), ]))

df = data.frame(sample = colnames(ncells), perc = Ex.perc)
gg = ggdensity(df, x = "perc", fill = "lightgray",
   add = "mean", rug = TRUE)
pdf(file.path(figures.path, "Ex_perc_density.pdf"), height = 4, width = 5)
print(gg)
dev.off()


ncells.freq = ncells.freq[, order(Ex.perc, decreasing = T)]
Ex.perc = (fast_column_sums(ncells.freq[grepl("^Ex", rownames(ncells.freq)) & !grepl("^Ex-NRGN", rownames(ncells.freq)), ]))

# SZ15 0.2570694
# SZ3 2.8237585
# SZ24 3.7128713
# SZ29 7.5571178

# mask = (fast_column_sums(ncells) > 1000) # (Ex.perc > 5)  & (Ex.perc < 80) & (fast_column_sums(ncells) > 500)
# 14 samples with less than 500 cells: CON32 CON43 CON47 CON50 SZ24  SZ25  SZ28  SZ31  SZ33  SZ34  SZ40  SZ41  SZ55  SZ60
# 1 Sample with more than 80% Ex: CON1
# 4 samples with less than 10% Ex: SZ37 SZ63 SZ64 SZ65


# > sort(colnames(ncells.freq)[fast_column_sums(ncells) < 500])
#  [1] "CON15" "CON17" "CON24" "CON35" "CON50" "CON55" "CON61" "CON67" "SZ39"  "SZ45"  "SZ48"  "SZ58"  "SZ6"   "SZ61" 
# > sort(colnames(ncells.freq)[Ex.perc > 80])
# [1] "SZ33"
# > sort(colnames(ncells.freq)[Ex.perc < 10])
# [1] "SZ15" "SZ24" "SZ29" "SZ3" 

mask = (Ex.perc >= 10) & (Ex.perc <= 80) & (fast_column_sums(ncells) >= 500)
# mask = !(colnames(ncells.freq) %in% c("SZ15"))
ncells.freq = ncells.freq[, mask]

# 
# 
# X = ncells.freq[, pb.logcounts$Cohort[match(colnames(ncells.freq), pb.logcounts$ID)] == "McLean"]
# df = reshape2::melt(X)
# colnames(df)=c("celltype", "sample", "freq")
# 
# df$celltype = factor(df$celltype, names(colors))
# # df$sample = droplevels(factor(df$sample, colnames(ncells.freq)[sample.perm]))
# 
# gg.ds1 = ggbarplot(df, "sample", "freq",
#   fill = "celltype", color = "black", palette = colors[levels(df$celltype)], xlab = "Individual", ylab = "Percentage") + theme(axis.text.x = element_text(face="bold", size=0, angle=90), axis.text.y = element_text(face="bold", size=12, angle=0), axis.title = element_text(face="bold", size=18, angle=0))
# 
# 
# pdf(file.path(figures.path, 'celltype_perc_McLean.pdf'), width = 10, height = 6)
# print(gg.ds1)
# dev.off()
# 
# 
# 
# X = ncells.freq[, pb.logcounts$Cohort[match(colnames(ncells.freq), pb.logcounts$ID)] != "McLean"]
# df = reshape2::melt(X)
# colnames(df)=c("celltype", "sample", "freq")
# 
# df$celltype = factor(df$celltype, names(colors))
# # df$sample = droplevels(factor(df$sample, colnames(ncells.freq)[sample.perm]))
# 
# gg.ds2 = ggbarplot(df, "sample", "freq",
#   fill = "celltype", color = "black", palette = colors[levels(df$celltype)], xlab = "Individual", ylab = "Percentage") + theme(axis.text.x = element_text(face="bold", size=0, angle=90), axis.text.y = element_text(face="bold", size=12, angle=0), axis.title = element_text(face="bold", size=18, angle=0))
# 
# 
# pdf(file.path(figures.path, 'celltype_perc_MtSinai.pdf'), width = 10, height = 6)
# print(gg.ds2)
# dev.off()
# 
# 
# pdf(file.path(figures.path, 'celltype_perc_combined.pdf'), width = 18, height = 6)
# gridExtra::grid.arrange(gg.ds1, gg.ds2, nrow = 1)
# dev.off()
# 
# 
# 
# 
# df$celltype = factor(df$celltype, names(colors))
# # df$sample = droplevels(factor(df$sample, colnames(ncells.freq)[sample.perm]))
# 
# gg.ds2 = ggbarplot(df, "sample", "freq",
#   fill = "celltype", color = "black", palette = colors[levels(df$celltype)], xlab = "Individual", ylab = "Percentage") + theme(axis.text.x = element_text(face="bold", size=0, angle=90), axis.text.y = element_text(face="bold", size=12, angle=0), axis.title = element_text(face="bold", size=18, angle=0))
# 
# 




X = ncells.freq
df = reshape2::melt(X)
colnames(df)=c("celltype", "sample", "freq")

df$celltype = factor(df$celltype, names(colors))

gg.combined = ggbarplot(df, "sample", "freq",
  fill = "celltype", color = "black", palette = colors[levels(df$celltype)], xlab = "Individual", ylab = "Percentage") + theme(axis.text.x = element_text(face="bold", size=8, angle=90,hjust=0.95,vjust=0.2), axis.text.y = element_text(face="bold", size=12, angle=0), axis.title = element_text(face="bold", size=18, angle=0))

pdf(file.path(figures.path, 'celltype_perc_joint.pdf'), width = 16, height = 6)
print(gg.combined)
dev.off()





X = ncells.freq[, pb.logcounts$Phenotype[match(colnames(ncells.freq), pb.logcounts$ID)] == "CON"]
df = reshape2::melt(X)
colnames(df)=c("celltype", "sample", "freq")

df$celltype = factor(df$celltype, names(colors))

gg.combined = ggbarplot(df, "sample", "freq",
  fill = "celltype", color = "black", palette = colors[levels(df$celltype)], xlab = "Individual", ylab = "Percentage") + theme(axis.text.x = element_text(face="bold", size=8, angle=90,hjust=0.95,vjust=0.2), axis.text.y = element_text(face="bold", size=12, angle=0), axis.title = element_text(face="bold", size=18, angle=0))

pdf(file.path(figures.path, 'celltype_perc_CON.pdf'), width = 12, height = 6)
print(gg.combined)
dev.off()



X = ncells.freq[, pb.logcounts$Phenotype[match(colnames(ncells.freq), pb.logcounts$ID)] == "SZ"]
df = reshape2::melt(X)
colnames(df)=c("celltype", "sample", "freq")

df$celltype = factor(df$celltype, names(colors))

gg.combined = ggbarplot(df, "sample", "freq",
  fill = "celltype", color = "black", palette = colors[levels(df$celltype)], xlab = "Individual", ylab = "Percentage") + theme(axis.text.x = element_text(face="bold", size=8, angle=90,hjust=0.95,vjust=0.2), axis.text.y = element_text(face="bold", size=12, angle=0), axis.title = element_text(face="bold", size=18, angle=0))

pdf(file.path(figures.path, 'celltype_perc_SZ.pdf'), width = 12, height = 6)
print(gg.combined)
dev.off()




```

# Plot cell type fraction stats per phenotype
```{r}
X.CON = apply(metadata(pb.logcounts)$n_cells[, pb.logcounts$Phenotype == "CON"], 2, as.numeric)
X.SZ = apply(metadata(pb.logcounts)$n_cells[, pb.logcounts$Phenotype == "SZ"], 2, as.numeric)
rownames(X.CON) = rownames(X.SZ) = names(assays(pb.logcounts))

# X.CON = apply(X.CON, 2, function(x) 100*x/sum(x))
# X.SZ = apply(X.SZ, 2, function(x) 100*x/sum(x))
# mu1 = apply(X.CON, 1, mean)
# mu2 = apply(X.SZ, 1, mean)
# sigmasq1 = apply(X.CON, 1, var)
# sigmasq2 = apply(X.SZ, 1, var)
# n1 = ncol(X.CON)
# n2 = ncol(X.SZ)

wilcox.out = presto::wilcoxauc(cbind(X.CON, X.SZ), c(rep("CON", ncol(X.CON)), rep("SZ", ncol(X.SZ))))
wilcox.out = wilcox.out[wilcox.out$group == "SZ", ]


Perc = t(apply(cbind(Matrix::rowMeans(X.CON), Matrix::rowMeans(X.SZ)), 1, function(x) 100*x / sum(x)))
colnames(Perc) = c("CON", "SZ")
Perc = Perc[order(Perc[, 1] - Perc[, 2], decreasing = T), ]
# df = Perc[wilcox.out$feature[order(-log10(wilcox.out$pval)*sign(wilcox.out$logFC))], ]
df = reshape2::melt(Perc)
colnames(df) = c("Celltype", "Phenotype", "Perc")
df$Celltype = factor(df$Celltype, names(colors))

gg = ggbarplot(df, "Celltype", "Perc", fill = "Phenotype", palette = c("#cccccc", "#888888"))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, colour = colors)) + xlab("Celltype") + ylab("Percentage")

pdf(file.path(figures.path, "Celltype_perc_per_phenotype.pdf"), height = 5, width = 8)
print(gg)
dev.off()




```



# Plot gene/umi statistics
## Per cell type
```{r}
umis = ACTIONet_summary$metadata$umi_count
mito.perc = ACTIONet_summary$metadata$mito_perc
genes = ACTIONet_summary$metadata$gene_counts
dataset = ACTIONet_summary$metadata$dataset
indiv = ACTIONet_summary$metadata$Individual
celltype = ACTIONet_summary$metadata$Labels
df = data.frame(celltype = celltype, umis = umis, genes = genes, mito = mito.perc, dataset = dataset, individual = indiv)  

df = df[df$individual %in% colnames(pb.logcounts)[mask == T], ]

df$celltype = factor(df$celltype, names(colors))

require(ggpubr)
gg.umis.1 = ggviolin(df[df$dataset == "McLean", ], "celltype", "umis", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("UMIs")+ theme(legend.position = "none")

png(file.path(figures.path, "McLean_UMIs.png"), width = 1200, height = 600, res = 150)
print(gg.umis.1)
dev.off()

gg.umis.2 = ggviolin(df[df$dataset == "MtSinai", ], "celltype", "umis", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("UMIs")+ theme(legend.position = "none")

png(file.path(figures.path, "MtSinai_UMIs.png"), width = 1200, height = 600, res = 150)
print(gg.umis.2)
dev.off()


require(ggpubr)
gg.genes.1 = ggviolin(df[df$dataset == "McLean", ], "celltype", "genes", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("genes")+ theme(legend.position = "none")

png(file.path(figures.path, "McLean_genes.png"), width = 1200, height = 600, res = 150)
print(gg.genes.1)
dev.off()

gg.genes.2 = ggviolin(df[df$dataset == "MtSinai", ], "celltype", "genes", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("genes")+ theme(legend.position = "none")

png(file.path(figures.path, "MtSinai_genes.png"), width = 1200, height = 600, res = 150)
print(gg.genes.2)
dev.off()

require(ggpubr)
gg.mito.1 = ggviolin(df[df$dataset == "McLean", ], "celltype", "mito", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("mito") + ylim(c(0, 10))+ theme(legend.position = "none")

png(file.path(figures.path,"McLean_mito.png"), width = 1200, height = 600, res = 150)
print(gg.mito.1)
dev.off()

gg.mito.2 = ggviolin(df[df$dataset == "MtSinai", ], "celltype", "mito", fill = "celltype", palette = colors,
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("mito") + ylim(c(0, 10))+ theme(legend.position = "none")

png(file.path(figures.path, "MtSinai_mito.png"), width = 1200, height = 600, res = 150)
print(gg.mito.2)
dev.off()

```

## Per sample
```{r}
df$individual = factor(as.character(df$individual), as.character(pb.logcounts$ID[order(pb.logcounts$umis, decreasing = T)]))

require(ggpubr)
gg.umis.1 = ggviolin(df[df$dataset == "McLean", ], "individual", "umis", fill = "individual", add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Individual") + ylab("UMIs")+ theme(legend.position = "none")


png(file.path(figures.path, "McLean_UMIs_per_sample.png"), width = 1200, height = 500, res = 150)
print(gg.umis.1)
dev.off()

gg.umis.2 = ggviolin(df[df$dataset == "MtSinai", ], "individual", "umis", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("UMIs")+ theme(legend.position = "none")

png(file.path(figures.path, "MtSinai_UMIs_per_sample.png"), width = 2100, height = 500, res = 150)
print(gg.umis.2)
dev.off()



df$individual = factor(as.character(df$individual), as.character(pb.logcounts$ID[order(pb.logcounts$genes, decreasing = T)]))

require(ggpubr)
gg.genes.1 = ggviolin(df[df$dataset == "McLean", ], "individual", "genes", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("genes")+ theme(legend.position = "none")

png(file.path(figures.path, "McLean_genes_per_sample.png"), width = 1200, height = 500, res = 150)
print(gg.genes.1)
dev.off()

gg.genes.2 = ggviolin(df[df$dataset == "MtSinai", ], "individual", "genes", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("genes")+ theme(legend.position = "none")

png(file.path(figures.path, "MtSinai_genes_per_sample.png"), width = 2100, height = 500, res = 150)
print(gg.genes.2)
dev.off()

df$individual = factor(as.character(df$individual), as.character(pb.logcounts$ID[order(pb.logcounts$mito_perc, decreasing = T)]))

require(ggpubr)
gg.mito.1 = ggviolin(df[df$dataset == "McLean", ], "individual", "mito", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("mito")+ theme(legend.position = "none") + ylim(c(0, 10))

png(file.path(figures.path,"McLean_mito_per_sample.png"), width = 1200, height = 500, res = 150)
print(gg.mito.1)
dev.off()

gg.mito.2 = ggviolin(df[df$dataset == "MtSinai", ], "individual", "mito", fill = "individual",
   add = "boxplot")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Celltype") + ylab("mito")+ theme(legend.position = "none") + ylim(c(0, 10))

png(file.path(figures.path, "MtSinai_mito_per_sample.png"), width = 2100, height = 500, res = 150)
print(gg.mito.2)
dev.off()

```


# Plot mappings of cell type annotations to other annotations
## Load markers
```{r}
data("curatedMarkers_human")
Layers = readRDS("~/results/input/Yao_layer_marker.RDS")
rdbu_fun = circlize::colorRamp2(c(-3, -1, 0, 1, 3), rev(pals::brewer.rdbu(9)[seq(1, 9, by = 2)]))
```

## Annotate
```{r}
# rownames(celltype.gene.spec)
# ct.layer.annot.1 = annotate.profile.using.markers(celltype.gene.spec, Layers)


X = as(do.call(cbind, lapply(Layers, function(gs) as.numeric(rownames(pb.logcounts) %in% gs))), "dgCMatrix")
rownames(X) = rownames(pb.logcounts)
Layer.annot = assess.geneset.enrichment.from.scores(celltype.gene.spec, X)

Z  = scale(t(Layer.annot$logPvals))
rownames(Z) = colnames(celltype.gene.spec)
# Z = scale(ct.layer.annot$Enrichment)
colnames(Z) = c(paste("L", c(1:6), sep = ""), "WM")

pdf(file.path(figures.path, "celltype_annotation_layers_v2.pdf"), width = 4)
Heatmap(Z, cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Layer", column_title = "Layer (Maynard et al.)")
dev.off()


# ct.mohammadi.annot = annotate.clusters.using.markers(ace, curatedMarkers_human$Brain$PFC$Mohammadi2020$marker.genes, specificity.slot.name = "Celltypes")

# Z = scale(ct.mohammadi.annot$Enrichment)
X = as(do.call(cbind, lapply(curatedMarkers_human$Brain$PFC$Mohammadi2020$marker.genes, function(gs) as.numeric(rownames(pb.logcounts) %in% gs))), "dgCMatrix")
rownames(X) = rownames(pb.logcounts)
Celltype.annot = assess.geneset.enrichment.from.scores(celltype.gene.spec, X)

Z  = scale(t(Celltype.annot$logPvals))
rownames(Z) = colnames(celltype.gene.spec)


M = as(t(MWM_hungarian(ct.mohammadi.annot$Enrichment)), "dgTMatrix")


pdf(file.path(figures.path, "celltype_annotation_mohammadi_markers_v2.pdf"), width = 6)
Heatmap(Z[, M@i+1], cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Celltypes", column_title = "Celltypes (Mohammadi et al.)")
dev.off()



ct.Velmeshev.annot = annotate.clusters.using.markers(ace, curatedMarkers_human$Brain$PFC$Velmeshev2019$marker.genes, specificity.slot.name = "Celltypes")

Z = scale(ct.Velmeshev.annot$Enrichment)

M = as(t(MWM_hungarian(ct.Velmeshev.annot$Enrichment)), "dgTMatrix")


pdf(file.path(figures.path, "celltype_annotation_velmeshev_markers.pdf"), width = 8)
Heatmap(Z[, M@i+1], cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Celltypes", column_title = "Celltypes (Velmeshev et al.)")
dev.off()


ct.Mathys.annot = annotate.clusters.using.markers(ace, curatedMarkers_human$Brain$PFC$MathysDavila2019$marker.genes, specificity.slot.name = "Celltypes")

Z = scale(ct.Mathys.annot$Enrichment)

M = as(t(MWM_hungarian(ct.Mathys.annot$Enrichment)), "dgTMatrix")


pdf(file.path(figures.path, "celltype_annotation_mathys_markers.pdf"), width = 8)
Heatmap(Z[, M@i+1], cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Celltypes", column_title = "Celltypes (Mathys and Davilla et al.)")
dev.off()



ct.Wang.annot = annotate.clusters.using.markers(ace, curatedMarkers_human$Brain$PFC$Wang2018$marker.genes, specificity.slot.name = "Celltypes")

Z = scale(ct.Wang.annot$Enrichment)

M = as(t(MWM_hungarian(ct.Wang.annot$Enrichment)), "dgTMatrix")


pdf(file.path(figures.path, "celltype_annotation_wang_markers.pdf"), width = 8)
Heatmap(Z[, M@i+1], cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Celltypes", column_title = "Celltypes (Wang et al.)")
dev.off()


ct.He.annot = annotate.clusters.using.markers(ace, curatedMarkers_human$Brain$Layers$marker.genes, specificity.slot.name = "Celltypes")

Z = scale(ct.He.annot$Enrichment)



pdf(file.path(figures.path, "celltype_annotation_He_markers.pdf"), width = 4)
Heatmap(Z, cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Layers", column_title = "Layers (He et al.)")
dev.off()








# X = as(do.call(cbind, lapply(Layers, function(gs) as.numeric(rownames(pb.logcounts) %in% gs))), "dgCMatrix")
# rownames(X) = rownames(pb.logcounts)
# Layer.annot = assess.geneset.enrichment.from.scores(ace$Celltypes_feature_specificity, X)
# 
# Annot = t(Layer.annot$logPvals)
# rownames(Annot) =  colnames(ace$Celltypes_feature_specificity)
# 
# Heatmap(scale(Annot[, -1]), cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"))

```
## Archetypes
```{r}
ct.arch.annot = annotate.archetypes.using.labels(t(ACTIONet_summary$H_unified), ACTIONet_summary$metadata$Labels)

perm = order(ct.arch.annot$Confidence, decreasing = T)
selected.cols = match(colnames(ct.arch.annot$Enrichment), ct.arch.annot$Label[perm])
selected.cols = selected.cols[!is.na(selected.cols)]
X = t(ct.arch.annot$Enrichment[perm, ])
X = X[, selected.cols]
Z = scale(X)

colnames(Z) = sapply(colnames(Z), function(str) str_split(str, "-")[[1]][[1]])

pdf(file.path(figures.path, "celltype_annotation_vs_archetypes_v3.pdf"), width = 7)
Heatmap(Z, cluster_rows = F, cluster_columns = F, rect_gp = gpar(col = "black"), row_names_side = "left", row_names_gp = gpar(col = colors), col = rdbu_fun, name = "Archetypes", column_title = "Archetypes")
dev.off()

```
# Plot cell states
```{r}
cellstates = c(7, 11, 17, 29)
cellstate.titles = c("A7-NRGN", "A11-Ex-SZ", "A17-SZTR", "A29-In-SZ")
subH = ACTIONet_summary$H_unified[, cellstates]

ggs = lapply(1:length(cellstates), function(i) {
  gg = plot.ACTIONet.gradient(ACTIONet_summary$ACTIONet2D, subH[, i], alpha_val = 0, point_size = 0.5, grad_palette = "magma") + ggtitle(cellstate.titles[[i]])
  
  return(gg)
})


png(file.path(figures.path, "ACTIONet_cellstates.png"), width = 800*3, height = 600*3, res = 150)
do.call(gridExtra::grid.arrange, c(lapply(ggs, ggplotify::as.grob), nrow = 2))
dev.off()

sapply(1:4, function(i) {
  png(file.path(figures.path, sprintf("ACTIONet_cellstates_%s.png", cellstate.titles[[i]])), width = 1600, height = 1200, res = 150)
  plot(ggs[[i]])
  dev.off()
})

```

# Plot cell types
```{r}
idx = setdiff(1:ncol(ACTIONet_summary$H_unified), cellstates)
subH = ACTIONet_summary$H_unified[, -cellstates]
Ll = ct.arch.annot$Label[-cellstates]
perm = order(match(Ll, names(colors)))
subH = subH[, perm]
Ll = Ll[perm]
idx = idx[perm]

ggs.archs.celltypes = lapply(1:length(Ll), function(i) {
  gg = plot.ACTIONet.gradient(ACTIONet_summary$ACTIONet2D, subH[, i], alpha_val = 0, point_size = 0.5, grad_palette = "magma") + ggtitle(sprintf("A%d-%s", idx[[i]], Ll[[i]])) +  theme(plot.title = element_text(color = colors[[Ll[[i]]]]))
  return(gg)
})


png(file.path(figures.path, "ACTIONet_archetypes_nonstate.png"), width = 800*7, height = 600*4, res = 150)
do.call(gridExtra::grid.arrange, c(lapply(ggs.archs.celltypes, ggplotify::as.grob), nrow = 4))
dev.off()
  
```



