---
title: "Perform DE analysis"
subtitle: "Step 1: Compute pseudobulk (PB) profiles"
output: html_notebook
---

# Setup
```{r include=FALSE}
require(ACTIONet)
require(stringr)
require(ComplexHeatmap)
require(muscat)


dataset.path = "~/results/datasets/"
results.path = "~/results"
figures.folder = "~/results/figures"


```

# Read ACTIONet and convert to an SCE object
```{r}
ace = readr::read_rds("~/results/ACTIONet_reunified.rds")
sce = as(ace, "SingleCellExperiment")

```

# Read metadata
```{r}
sample.metadata = readr::read_rds(file.path(dataset.path, "sample_metadata.rds"))
ACTIONet_summary = readr::read_rds(file.path(dataset.path, "ACTIONet_summary.rds"))

```

# Use muscat to compute PB profiles
```{r}
sce$cluster_id = ACTIONet_summary$metadata$Labels
sce$group_id = ACTIONet_summary$metadata$Phenotype
sce$sample_id = ACTIONet_summary$metadata$Individual

library(muscat)
sce$id <- paste0(sce$Phenotype, sce$sample_id)
(sce <- prepSCE(sce,
    kid = "cluster_id", # subpopulation assignments
    gid = "group_id",  # group IDs (ctrl/stim)
    sid = "sample_id",   # sample IDs (ctrl/stim.1234)
    drop = TRUE))  # drop all other colData columns

pb.logcounts <- aggregateData(sce,
    assay = "logcounts", fun = "mean",
    by = c("cluster_id", "sample_id"))


colData(pb.logcounts) = cbind(colData(pb.logcounts), sample.metadata[colnames(pb.logcounts), ])
readr::write_rds(pb.logcounts, file = file.path("~/results/datasets/PB_mean_logcounts_final.RDS")

```

```{r}
pb.logcounts.old = readr::read_rds(file.path(dataset.path, "PB_sce_final.rds"))
pb.logcounts.old = readr::read_rds(file.path(dataset.path, "PB_sce_final.rds"))


metas = readr::read_rds("~/results/datasets/meta_data_final.rds")


metas2 =data.frame(colData(pb.logcounts.old))

readr::write_rds(pb.logcounts, file = "~/results/PB_mean_logcounts_renormalized_with_subtypes_annotated.RDS")



sample.metadata = data.frame(colData(pb.logcounts))
sample.metadata = sample.metadata[, -c(29:32, 42:44)]



```

```{r}
ACTIONet_summary = readr::read_rds(file.path(dataset.path, "ACTIONet_summary_filtered_individuals.rds"))

pb.logcounts = readr::read_rds(file.path(dataset.path, "PB_sce_final.rds"))

color.df = readRDS(file.path(dataset.path, "celltype_colors.rds"))

colors = color.df$color
names(colors) = color.df$celltype

# int_colData(pb.logcounts)$n_cells = t(metadata(pb.logcounts)$n_cells)
# metadata(pb.logcounts) = metadata(pb.logcounts)[c(1:2)]
# ncells = t(apply(int_colData(pb.logcounts)$n_cells, 2, as.numeric))

ncells = apply(metadata(pb.logcounts)$n_cells, 2, as.numeric)





```

# Prefilter outlier samples wrt ExNero %, or samples that have less than 500 cells (pruning from 140 -> 121 samples)
```{r}
rownames(ncells) = names(assays(pb.logcounts))
cs = Matrix::colSums(ncells)
ncells.freq = 100*scale(ncells, center = F, scale = cs)
Ex.perc = (fast_column_sums(ncells.freq[grepl("^Ex", rownames(ncells.freq)) & !grepl("^Ex-NRGN", rownames(ncells.freq)), ]))

ncells.freq = ncells.freq[, order(Ex.perc, decreasing = T)]
Ex.perc = (fast_column_sums(ncells.freq[grepl("^Ex", rownames(ncells.freq)) & !grepl("^Ex-NRGN", rownames(ncells.freq)), ]))


mask = (Ex.perc > 10)  & (Ex.perc < 80) & (fast_column_sums(ncells) > 500)

# Samples that are depleted in Ex.perc:: 
# # SZ15 0.2570694
# # SZ3 2.8237585
# # SZ24 3.7128713
# # SZ29 7.5571178

pb.logcounts.filtered = pb.logcounts[, mask]
metadata(pb.logcounts.filtered)$n_cells = metadata(pb.logcounts.filtered)$n_cells[, mask]


```

# Performing cohort-specific DE
```{r}
form = ~ Phenotype + Batch + Gender + Age + PMI + Benzodiazepines + Anticonvulsants + AntipsychTyp + AntipsychAtyp + Antidepress

resDE = lapply( levels(pb.logcounts.filtered$Cohort), function(chrt){

	keep.ids = colnames(pb.logcounts.filtered)[pb.logcounts.filtered$Cohort == chrt]

	pb.logcounts.filtered_sub = pb.logcounts.filtered[,keep.ids]
	metadata(pb.logcounts.filtered_sub)$n_cells = metadata(pb.logcounts.filtered_sub)$n_cells[,keep.ids]

	design.mat <- model.matrix(form, data = droplevels(colData(pb.logcounts.filtered_sub)))

	colnames(design.mat)[1] = c("Intercept")

	contrast.mat <- makeContrasts(contrasts = "PhenotypeSZ", levels = design.mat)

	pbDS(pb.logcounts.filtered_sub, method = "limma-trend", min_cells = 5, design = design.mat, contrast =  contrast.mat, filter = "gene")
})
names(resDE) = levels(colData(pb.logcounts.filtered)$Cohort)

readr::write_rds(resDE, file.path(dataset.path, "Cohort_specific_DE_results.rds"))



```

# Read bulk DE
```{r}
CMC.DE = read.table("~/results/input/CMC_MSSM-Penn-Pitt_DLPFC_mRNA_IlluminaHiSeq2500_gene-adjustedSVA-differentialExpression-includeAncestry-DxSCZ-DE.tsv", header = T)
PEC.DE = read.table("~/results/input/Full_DE_table.csv", header = T)

common.genes = intersect(rownames(pb.logcounts), intersect(CMC.DE$MAPPED_genes, PEC.DE$gene_name))

CMC.tstat = CMC.DE$t[match(common.genes, CMC.DE$MAPPED_genes)]
PEC.tstat = PEC.DE$SCZ.t.value[match(common.genes, PEC.DE$gene_name)]

names(PEC.tstat) = names(CMC.tstat) = common.genes
pval = cor.test(CMC.tstat, PEC.tstat, method = "spearman")$p.value

McLean.PEC.cor.pvals = sapply(resDE$McLean$table$PhenotypeSZ, function(DF) {
  tstats = DF$t
  names(tstats) = DF$gene
  cg = intersect(common.genes, DF$gene)
  x = -log10(cor.test(tstats[cg], PEC.tstat[cg], method = "spearman")$p.value)
  
  return(x)
})


MtSinai.PEC.cor.pvals = sapply(resDE$MtSinai$table$PhenotypeSZ, function(DF) {
  tstats = DF$t
  names(tstats) = DF$gene
  cg = intersect(common.genes, DF$gene)
  x = -log10(cor.test(tstats[cg], PEC.tstat[cg], method = "spearman")$p.value)
  
  return(x)
})


McLean.CMC.cor.pvals = sapply(resDE$McLean$table$PhenotypeSZ, function(DF) {
  tstats = DF$t
  names(tstats) = DF$gene
  cg = intersect(common.genes, DF$gene)
  x = -log10(cor.test(tstats[cg], CMC.tstat[cg], method = "spearman")$p.value)
  
  return(x)
})


MtSinai.CMC.cor.pvals = sapply(resDE$MtSinai$table$PhenotypeSZ, function(DF) {
  tstats = DF$t
  names(tstats) = DF$gene
  cg = intersect(common.genes, DF$gene)
  x = -log10(cor.test(tstats[cg], CMC.tstat[cg], method = "spearman")$p.value)
  
  return(x)
})

df = data.frame(Celltypes = names(resDE$MtSinai$table$PhenotypeSZ), McLean.PEC = McLean.PEC.cor.pvals[names(resDE$MtSinai$table$PhenotypeSZ)], MtSinai.PEC = MtSinai.PEC.cor.pvals[names(resDE$MtSinai$table$PhenotypeSZ)], McLean.CMC = McLean.CMC.cor.pvals[names(resDE$MtSinai$table$PhenotypeSZ)], MtSinai.CMC = MtSinai.CMC.cor.pvals[names(resDE$MtSinai$table$PhenotypeSZ)])
df$Combined = combine.logPvals(t(as.matrix(df[, -1])))
df = df[order(df$Combined, decreasing = T), ]

df2 = reshape2::melt(df)
colnames(df2) = c("Celltype", "Test", "Enrichment")
df2$Test = factor(as.character(df2$Test), c("Combined", "McLean.PEC", "MtSinai.PEC", "McLean.CMC", "MtSinai.CMC"))
df2$Celltype = factor(df2$Celltype, df$Celltypes[order(df$Combined, decreasing = T)])

require(ggpubr)
gg =ggbarplot(df2, "Celltype", "Enrichment", fill = "Test", color = "black", palette = rev(pals::brewer.paired(5)),
  position = position_dodge(0.9), xlab = "Celltype", ylab = "Enrichment")+ theme(axis.text.x = element_text(face="bold", size=8, angle=90,hjust=0.95,vjust=0.2, color = colors[levels(df2$Celltype)]), axis.text.y = element_text(face="bold", size=12, angle=0), axis.title = element_text(face="bold", size=18, angle=0))

pdf(file.path(figures.folder, "DE_vs_bulk_v0.pdf"))
plot(gg)
dev.off()




```

```{r}
X = log1p(ACTIONet_summary$unified_feature_specificity[common.genes, ])
arch.SZ.CC = cor(X, abs(cbind(CMC.tstat, PEC.tstat)), method = "pearson")

```

# Prefiltering individual DE results before combining them
Only keep genes that have 1) consistent direction of dysregulation across both datasets, and 2) at least 1 sd away from zero on the logFC scale

```{r}
celltypes = intersect(names(resDE$McLean$table$PhenotypeSZ), names(resDE$MtSinai$table$PhenotypeSZ))

filtered.tables = lapply(celltypes, function(celltype) {
  tbl1 = resDE[[1]]$table$PhenotypeSZ[[celltype]]
  tbl2 = resDE[[2]]$table$PhenotypeSZ[[celltype]]
  
  genes = intersect(tbl1$gene[1 <= abs(tbl1$logFC/sd(tbl1$logFC))], tbl2$gene[1 <= abs(tbl2$logFC / sd(tbl2$logFC))])
  
  tbl1 = tbl1[match(genes, tbl1$gene), ]
  tbl2 = tbl2[match(genes, tbl2$gene), ]
  
  mask = sign(tbl1$logFC)*sign(tbl2$logFC) > 0
  tbl1 = tbl1[mask, ]
  tbl2 = tbl2[mask, ]
  
  tbls = list(McClean = tbl1, MtSinai = tbl2) 
  tbls = lapply( tbls, function(tab){
    tab$se = tab$logFC / tab$t
    tab
  })
  
  return(tbls)
})
names(filtered.tables) = celltypes

readr::write_rds(filtered.tables, file.path(dataset.path, "individual_diff_results_filtered.rds"))

```



```{r}
# filtered.tables.old = readRDS("~/PFC_v3/individual_diff_results_filtered.rds")
# 
# i = 1
X1 = filtered.tables.old$`Ex-CC_THEMIS`
X2 = filtered.tables$`Ex-L6_CC_SEMA3A`


cor(X1$McClean$t, X2$McClean$t)

resDE.old = readr::read_rds("~/PFC_v3/filtered_resDE.rds")


resDE.old$McLean$table$PhenotypeSZ$`Ex-CC_THEMIS`$t

```


