---
title: "Perform DE analysis"
output: html_notebook
---

# Setup
```{r include=FALSE}
require(ACTIONet)
require(stringr)
require(ComplexHeatmap)

require(muscat)
require(edgeR)
require(limma)

dataset.path = "~/results/datasets/"
results.path = "~/results"
figures.folder = "~/results/figures"


```





```{r}
ACTIONet_summary = readr::read_rds(file.path(dataset.path, "ACTIONet_summary_filtered_individuals.rds"))

pb.logcounts = readr::read_rds(file.path(dataset.path, "PB_mean_logcounts_final.RDS"))

color.df = readRDS(file.path(dataset.path, "celltype_colors.rds"))

# color.df[c(22:24), ] = color.df[c(24, 23, 22), ]

colors = color.df$color
names(colors) = color.df$celltype


ncells = sapply(int_colData(pb.logcounts)$n_cells, as.numeric)
rownames(ncells) = names(assays(pb.logcounts))

```






# Prefilter outlier samples wrt ExNero %, or samples that have less than 500 cells (pruning from 140 -> 121 samples)
```{r}
cs = Matrix::colSums(ncells)
ncells.freq = 100*scale(ncells, center = F, scale = cs)
Ex.perc = (fast_column_sums(ncells.freq[grepl("^Ex", rownames(ncells.freq)) & !grepl("^Ex-NRGN", rownames(ncells.freq)), ]))

# mask = !(pb.logcounts$ID %in% c("SZ15"))
mask = (Ex.perc >= 10) & (Ex.perc <= 80) #& (fast_column_sums(ncells) >= 500)
# mask = (Ex.perc >= 1) & (fast_column_sums(ncells) >= 500)
# mask = (Ex.perc >= 1)
pb.logcounts.filtered = pb.logcounts [, mask]


ff = interaction(ACTIONet_summary$metadata$Individual, ACTIONet_summary$metadata$Labels)
umi.tbl = split(ACTIONet_summary$metadata$umi_count, ff)
mean.celltype.umi = sapply(umi.tbl, function(x) {
  if(length(x) == 0) {
    return(0)
  } else {
    return(mean(x))
  }
})
mean.celltype.umi.mat = matrix(mean.celltype.umi, nrow = length(levels(ACTIONet_summary$metadata$Individual)))


```

# Performing cohort-specific DE
```{r}
# int_colData(pb.logcounts.filtered)$n_cells = t(metadata(pb.logcounts.filtered)$n_cells)
# metadata(pb.logcounts.filtered) = metadata(pb.logcounts.filtered)[c(1:2)]

#  [1] "Cohort"          "ID"              "Internal_ID"     "CMC_ID"          "Phenotype"       "Batch"           "HTO"            
#  [8] "Gender"          "Age"             "PMI"             "EUR_Ancestry"    "EAS_Ancestry"    "AMR_Ancestry"    "SAS_Ancestry"   
# [15] "AFR_Ancestry"    "Benzodiazepines" "Anticonvulsants" "AntipsychTyp"    "AntipsychAtyp"   "Antidepress"     "Lithium"        
# [22] "MedPC1"          "MedPC2"          "MedPC3"          "MedPC4"          "MedPC5"          "MedPC6"          "Cells"          
# [29] "Density"         "SZTR.mean"       "A7.signature"    "A11.signature"   "A17.signature"   "A29.signature"   "PRS"            
# [36] "POP.EL3SD"       "genes"           "umis"            "mito_perc"       "TPS.Ex"          "TPS.Neuro"       "TPS.All" 
# form = ~ Phenotype + Batch + Gender + Age + PMI + Benzodiazepines + Anticonvulsants + AntipsychTyp + AntipsychAtyp + Antidepress
# form = ~ Phenotype + Batch + Gender + Age + PMI + Benzodiazepines + Anticonvulsants + AntipsychTyp + AntipsychAtyp + Antidepress + EUR_Ancestry + EAS_Ancestry + AMR_Ancestry + SAS_Ancestry + AFR_Ancestry

# pb.logcounts.filtered$logumis = log1p(pb.logcounts.filtered$umis)

pb.logcounts.filtered$SampleQuality = scale(log1p(pb.logcounts.filtered$umis))
form = ~ Phenotype + Batch + PMI + Gender + Age + Benzodiazepines + Anticonvulsants + AntipsychTyp + AntipsychAtyp + Antidepress + SampleQuality

resDE = lapply( levels(pb.logcounts.filtered$Cohort), function(chrt){

	keep.ids = colnames(pb.logcounts.filtered)[pb.logcounts.filtered$Cohort == chrt]

	pb.logcounts.filtered_sub = pb.logcounts.filtered[,keep.ids]
	# metadata(pb.logcounts.filtered_sub)$n_cells = metadata(pb.logcounts.filtered_sub)$n_cells[,keep.ids]
  sample.metadata = droplevels(data.frame(colData(pb.logcounts.filtered_sub)))
	design.mat <- model.matrix(form, data = sample.metadata)

# #   sample.quality.mat = apply(mean.celltype.umi.mat[pb.logcounts.filtered_sub$ID, ], 2, function(x) log2(x / mean(x)))
# # 	sample.quality.mat[sample.quality.mat == -Inf] = min(sample.quality.mat[!is.infinite(sample.quality.mat)])
#   sample.quality.mat = log1p(apply(mean.celltype.umi.mat[pb.logcounts.filtered_sub$ID, ], 2, function(x) x))
#   
#   
# 	sample.quality.mat[sample.quality.mat == 0] = NA
#   sample.quality.mat.z = apply(sample.quality.mat, 2, function(x) x / sum(x, na.rm = T))
# 
#     sample.quality = as.numeric(scale(apply(sample.quality.mat.z, 1, function(x) mean(x, na.rm = T)	)))

#   
#   
# 	sample.quality.mat[sample.quality.mat == 0] = NA
#   sample.quality.mat.z = apply(sample.quality.mat, 2, function(x) (x - median(x, na.rm = T))/mad(x, na.rm = T))
# 	sample.quality.mat.z[sample.quality.mat.z < -3] = -3
# 	sample.quality.mat.z[sample.quality.mat.z > 3] = 3
# 
#   # sample.quality.mat = apply(sample.quality.mat, 2, function(x) x / sum(x))
# 
#   
#   design.mat = cbind(design.mat, sample.quality)
#   colnames(design.mat)[[ncol(design.mat)]] = "SampleQuality"
	

	# colnames(design.mat) = c("Intercept", "PhenotypeSZ", paste("V", 3:ncol(design.mat), sep = ""))
	colnames(design.mat)[1] = c("Intercept")

	contrast.mat <- makeContrasts(contrasts = "PhenotypeSZ", levels = design.mat)

	df = pbDS(pb.logcounts.filtered_sub, method = "limma-trend", min_cells = 5, design = design.mat, contrast =  contrast.mat, filter = "both")
	
})
names(resDE) = levels(colData(pb.logcounts.filtered)$Cohort)

readr::write_rds(resDE, file.path(dataset.path, "Cohort_specific_DE_results_final_with_logumi_baseline_full_filter_extended_cell_filtering.rds"))

# resDE = readr::read_rds("~/PFC_v3/")
# lst_meta = readr::read_rds("~/PFC_v3/filtered_resDE.rds")


```


## Export as excel tables
```{r}
for(ds in 1:2) {

  library(openxlsx)
  Up.wb <- createWorkbook()
  for(i in 1:length(resDE[[ds]]$table$PhenotypeSZ)) {
    res = resDE[[ds]]$table$PhenotypeSZ[[i]]
    res = res[res$logFC > 0, ]
    res = cbind(data.frame(Gene = rownames(res)), res)
    res = res[order(res$t, decreasing = T), ]
  
    n = names(resDE[[ds]]$table$PhenotypeSZ)[[i]] #str_replace(arch.names[arch.order[i]], "/", "-")
    
    addWorksheet(wb=Up.wb, sheetName = n)
    writeData(Up.wb, sheet = n, res) 
  
  }
  saveWorkbook(Up.wb, sprintf(file.path(dataset.path, "DE_genes_up_%s_with_logumi_baseline_full_filter_extended_cell_filtering_fullset.xlsx"), names(resDE)[[ds]]), overwrite = TRUE)
  
  
  library(openxlsx)
  Down.wb <- createWorkbook()
  for(i in 1:length(resDE[[ds]]$table$PhenotypeSZ)) {
    res = resDE[[ds]]$table$PhenotypeSZ[[i]]
    res = res[res$logFC < 0, ]
    res = cbind(data.frame(Gene = rownames(res)), res)
    res = res[order(res$t, decreasing = F), ]
    
    n = names(resDE[[ds]]$table$PhenotypeSZ)[[i]] #str_replace(arch.names[arch.order[i]], "/", "-")
    
    addWorksheet(wb=Down.wb, sheetName = n)
    writeData(Down.wb, sheet = n, res) 
  
  }
  saveWorkbook(Down.wb, sprintf(file.path(dataset.path, "DE_genes_down_%s_with_logumi_baseline_full_filter_extended_cell_filtering_fullset.xlsx"), names(resDE)[[ds]]), overwrite = TRUE)
}

```



# Prefiltering individual DE results before combining them
Only keep genes that have 1) consistent direction of dysregulation across both datasets, and 2) at least 1 sd away from zero on the logFC scale

```{r}
celltypes = intersect(names(resDE$McLean$table$PhenotypeSZ), names(resDE$MtSinai$table$PhenotypeSZ))

filtered.tables = lapply(celltypes, function(celltype) {
  tbl1 = resDE[[1]]$table$PhenotypeSZ[[celltype]]
  tbl2 = resDE[[2]]$table$PhenotypeSZ[[celltype]]
  
  genes = intersect(tbl1$gene[1 <= abs(tbl1$logFC/sd(tbl1$logFC))], tbl2$gene[1 <= abs(tbl2$logFC / sd(tbl2$logFC))])
  # genes = intersect(tbl1$gene, tbl2$gene)
  
  tbl1 = tbl1[match(genes, tbl1$gene), ]
  tbl2 = tbl2[match(genes, tbl2$gene), ]
  
  mask = sign(tbl1$logFC)*sign(tbl2$logFC) > 0
  tbl1 = tbl1[mask, ]
  tbl2 = tbl2[mask, ]
  
  tbls = list(McClean = tbl1, MtSinai = tbl2) 
  tbls = lapply( tbls, function(tab){
    tab$se = tab$logFC / tab$t
    tab
  })
  
  return(tbls)
})
names(filtered.tables) = celltypes

readr::write_rds(filtered.tables, file.path(dataset.path, "individual_diff_results_filtered_full_with_logumi_baseline_full_filter_extended_cell_filtering.rds"))

```

## Export as excel tables
```{r}
for(ds in 1:2) {

  library(openxlsx)
  Up.wb <- createWorkbook()
  for(i in 1:length(filtered.tables)) {
    res = filtered.tables[[i]][[ds]]
    res = res[res$logFC > 0, ]
    res = cbind(data.frame(Gene = rownames(res)), res)
    res = res[order(res$t, decreasing = T), ]
  
    n = names(filtered.tables)[[i]] #str_replace(arch.names[arch.order[i]], "/", "-")
    
    addWorksheet(wb=Up.wb, sheetName = n)
    writeData(Up.wb, sheet = n, res) 
  
  }
  saveWorkbook(Up.wb, sprintf(file.path(dataset.path, "DE_genes_up_%s_with_logumi_baseline_full_filter_extended_cell_filtering.xlsx"), names(filtered.tables[[i]])[[ds]]), overwrite = TRUE)
  
  
  library(openxlsx)
  Down.wb <- createWorkbook()
  for(i in 1:length(filtered.tables)) {
    res = filtered.tables[[i]][[ds]]
    res = res[res$logFC < 0, ]
    res = cbind(data.frame(Gene = rownames(res)), res)
    res = res[order(res$t, decreasing = F), ]
    
    n = names(filtered.tables)[[i]] #str_replace(arch.names[arch.order[i]], "/", "-")
    
    addWorksheet(wb=Down.wb, sheetName = n)
    writeData(Down.wb, sheet = n, res) 
  
  }
  saveWorkbook(Down.wb, sprintf(file.path(dataset.path, "DE_genes_down_%s_with_logumi_baseline_full_filter_extended_cell_filtering.xlsx"), names(filtered.tables[[i]])[[ds]]), overwrite = TRUE)
}

```


```{r}
combined.analysis.tables = lapply(names(filtered.tables), function(celltype) {
  print(celltype)
  tbls = filtered.tables[[celltype]]
  
  gene.tbls = lapply(1:nrow(tbls[[1]]), function(i) {
    dfs = lapply(1:length(tbls), function(k) tbls[[k]][i, ])
    df = do.call("rbind", dfs)
  })
  names(gene.tbls) = tbls[[1]]$gene
    
  combined.analysis.tbl = do.call(rbind, lapply(names(gene.tbls), function(gene){
    x = suppressWarnings(metafor::rma(yi=logFC, sei=se, data = gene.tbls[[gene]], method="FE"))
    combined.tbl = data.frame( gene = gene, 
        logFC     = x$beta,
        se        = x$se,
        tstat = x$zval,
        P.Value   = x$pval)
    return(combined.tbl)
  }))
  rownames(combined.analysis.tbl) = names(gene.tbls)
  
  combined.analysis.tbl = combined.analysis.tbl[order(combined.analysis.tbl$P.Value), ]
  
  return(combined.analysis.tbl)
})
names(combined.analysis.tables) = names(filtered.tables)

DF = do.call(rbind, combined.analysis.tables)
DF$adj.P.Val = p.adjust(DF$P.Value, "fdr")
ff = factor(unlist(lapply(names(combined.analysis.tables), function(celltype) rep(celltype, nrow(combined.analysis.tables[[celltype]])))), names(filtered.tables))
combined.analysis.tables = split(DF, ff)

readr::write_rds(combined.analysis.tables, file.path(dataset.path, "meta_analysis_diff_results_full_with_logumi_baseline_full_filter_extended_cell_filtering.rds"))

```



```{r eval=FALSE, include=FALSE}
resDE = readr::read_rds(file.path(dataset.path, "filtered_resDE_with_logumi_baseline_full_filter_extended_cell_filtering.rds"))

filtered.tables = readr::read_rds(file.path(dataset.path, "individual_diff_results_filtered_full_with_logumi_baseline_full_filter_extended_cell_filtering.rds"))

combined.analysis.tables = readr::read_rds(file.path(dataset.path, "meta_analysis_diff_results_with_logumi_baseline_full_filter_extended_cell_filtering.rds"))

```

```{r}
  library(openxlsx)
  Up.wb <- createWorkbook()
  for(i in 1:length(combined.analysis.tables)) {
    res = combined.analysis.tables[[i]]
    res = res[(res$logFC > 0.1) & (res$P.Value <= 0.05), ]
    res = res[order(res$t, decreasing = T), ]
  
    n = names(combined.analysis.tables)[[i]] #str_replace(arch.names[arch.order[i]], "/", "-")
    
    addWorksheet(wb=Up.wb, sheetName = n)
    writeData(Up.wb, sheet = n, res) 
  
  }
  saveWorkbook(Up.wb, sprintf(file.path(dataset.path, "DE_genes_up_%s_with_logumi_baseline_full_filter_extended_cell_filtering.xlsx"), "combined"), overwrite = TRUE)
  
  
  library(openxlsx)
  Down.wb <- createWorkbook()
  for(i in 1:length(combined.analysis.tables)) {
    res = combined.analysis.tables[[i]]
    res = res[(res$logFC < -0.1) & (res$P.Value <= 0.05), ]
    res = res[order(res$t, decreasing = F), ]
    
    n = names(combined.analysis.tables)[[i]] #str_replace(arch.names[arch.order[i]], "/", "-")
    
    addWorksheet(wb=Down.wb, sheetName = n)
    writeData(Down.wb, sheet = n, res) 
  
  }
  saveWorkbook(Down.wb, sprintf(file.path(dataset.path, "DE_genes_down_%s_with_logumi_baseline_full_filter_extended_cell_filtering.xlsx"), "combined"), overwrite = TRUE)
```


```{r}
DE.sc = matrix(0, nrow(pb.logcounts), length(combined.analysis.tables))
tstats = matrix(0, nrow(pb.logcounts), length(combined.analysis.tables))
logFC = matrix(0, nrow(pb.logcounts), length(combined.analysis.tables))
logPvals = matrix(0, nrow(pb.logcounts), length(combined.analysis.tables))
rownames(DE.sc) = rownames(tstats) = rownames(logFC) = rownames(logPvals) = rownames(pb.logcounts)
colnames(DE.sc) = colnames(tstats) = colnames(logFC) = colnames(logPvals) = names(combined.analysis.tables)

limma_trend_mean.scores = matrix(0, nrow(pb.logcounts), length(combined.analysis.tables))
Up.genes = vector("list", length(combined.analysis.tables))
Down.genes = vector("list", length(combined.analysis.tables))
rownames(limma_trend_mean.scores) = rownames(pb.logcounts)
names(Up.genes) = names(Down.genes) = colnames(limma_trend_mean.scores) = names(combined.analysis.tables)
for(i in 1:length(combined.analysis.tables)) {
	print(i)
	
	tbl = combined.analysis.tables[[i]]

	tstats[tbl$gene, names(combined.analysis.tables)[[i]]] = tbl$tstat
	logFC[tbl$gene, names(combined.analysis.tables)[[i]]] = tbl$logFC
	logPvals[tbl$gene, names(combined.analysis.tables)[[i]]] = -log10(tbl$adj.P.Val)
	
	DE.sc[tbl$gene, names(combined.analysis.tables)[[i]]] = tbl$tstat
	
}
limma_trend_mean.scores[is.na(limma_trend_mean.scores)] = 0


Up.genes = lapply(combined.analysis.tables, function(combined.analysis.tbl) {
  combined.analysis.tbl$gene[(combined.analysis.tbl$logFC > 0.1) & (combined.analysis.tbl$adj.P.Val < 0.05)]
})
Down.genes = lapply(combined.analysis.tables, function(combined.analysis.tbl) {
  combined.analysis.tbl$gene[(combined.analysis.tbl$logFC < -0.1) & (combined.analysis.tbl$adj.P.Val < 0.05)]
})

DE.new = list(DE.sc = DE.sc, tstats = tstats, logFC = logFC, logPvals = logPvals, Up.genes = Up.genes, Down.genes = Down.genes)

saveRDS(DE.new, file.path(dataset.path, "DE_genes_pseudobulk_final_with_logumi_baseline_full_filter_extended_cell_filtering.rds"))



```


