---
title: "Fig3: H-MAGMA analysis"
output: html_notebook
---

# Setup
```{r include=FALSE}
library(org.Hs.eg.db)

dataset.path = "~/results/datasets/"
results.path = "~/results"
figures.folder = "~/results/figures"

MAGMA.path = "~/magma"


```

# Load DE genes
```{r include=FALSE}
DE.new = readRDS(file.path(dataset.path, "DE_genes_pseudobulk_final_with_umi.rds"))

Up.genes = DE.new$Up.genes
Down.genes = DE.new$Down.genes

color.df = readRDS(file.path(dataset.path, "celltype_colors.rds"))

colors = color.df$color
names(colors) = color.df$celltype

```



```{r}
# Up.genes = sapply(resDE$McLean$table$PhenotypeSZ, function(df) {
#   mask = (df$p_val < 0.05) & (df$logFC > 0.1)
#   df$gene[mask]
# })
# Down.genes = sapply(resDE$McLean$table$PhenotypeSZ, function(df) {
#   mask = (df$p_val < 0.05) & (df$logFC < -0.1)
#   df$gene[mask]
# })





```



# Convert DE genes to ENSG and export
```{r}
require(org.Hs.eg.db)

suppressWarnings(ids <- AnnotationDbi::mapIds(org.Hs.eg.db, keys = row.names(DE.new$DE.sc), keytype = "SYMBOL", column = "ENSEMBL", multiVals = "first"))
ids[is.na(ids)] = ""

Up.genes.ENSG = sapply(Up.genes, function(gs) {
  setdiff(sort(unique(ids[match(gs, rownames(pb.logcounts))])), "")
})
Up.genes.ENSG.df = reshape2::melt(Up.genes.ENSG)
Up.genes.ENSG.df = Up.genes.ENSG.df[, c(2, 1)]
Up.genes.ENSG.df$L1 = factor(Up.genes.ENSG.df$L1, names(Up.genes))
Up.genes.ENSG.df = Up.genes.ENSG.df[order(Up.genes.ENSG.df$L1), ]
write.table(Up.genes.ENSG.df, file.path(MAGMA.path, "genesets", "Up_genes_ENSG.tsv"), sep = "\t", col.names = F, row.names = F, quote = F)

Down.genes.ENSG = sapply(Down.genes, function(gs) {
  setdiff(sort(unique(ids[match(gs, rownames(pb.logcounts))])), "")
})
Down.genes.ENSG.df = reshape2::melt(Down.genes.ENSG)
Down.genes.ENSG.df = Down.genes.ENSG.df[, c(2, 1)]
Down.genes.ENSG.df$L1 = factor(Down.genes.ENSG.df$L1, names(Down.genes))
Down.genes.ENSG.df = Down.genes.ENSG.df[order(Down.genes.ENSG.df$L1), ]

write.table(Down.genes.ENSG.df, file.path(MAGMA.path, "genesets", "Down_genes_ENSG.tsv"), sep = "\t", col.names = F, row.names = F, quote = F)


UpAndDown.genes.ENSG.df = rbind(Up.genes.ENSG.df, Down.genes.ENSG.df)
UpAndDown.genes.ENSG.df$L1 = factor(UpAndDown.genes.ENSG.df$L1, names(Up.genes))
UpAndDown.genes.ENSG.df = UpAndDown.genes.ENSG.df[order(UpAndDown.genes.ENSG.df$L1), ]

write.table(UpAndDown.genes.ENSG.df, file.path(MAGMA.path, "genesets", "UpAndDown_genes_ENSG.tsv"), sep = "\t", col.names = F, row.names = F, quote = F)


```





# Run H-MAGMA
We run HMAGMA using  **for f in `ls hmagma/`; do (./magma --gene-results ./hmagma/$f/$f.genes.raw --set-annot genesets/UpAndDown_genes_ENSG.tsv col=2,1 --out $f\_UpAndDown &); done`**, from inside the magma folder, where hmagma are prestored the gene models scores.

# Load H-MAGMA results

```{r}
dl = list.dirs(file.path(MAGMA.path, "psych_arch/"), full.names = F, recursive = F)

HMAGMA.Pvals = vector("list", 1)
names(HMAGMA.Pvals) = c("UpAndDown")
for (d in names(HMAGMA.Pvals)) {
  HMAGMA.Pvals[[d]] = matrix(1, nrow = 20, ncol = length(dl))
  colnames(HMAGMA.Pvals[[d]]) = dl
  rownames(HMAGMA.Pvals[[d]]) = names(Up.genes)
  for(cond in dl) {
    print(cond)
    file.name = sprintf("%s/out/%s_%s.gsa.out", MAGMA.path, cond, d)
    lines = readLines(con <- file(file.name))
    lines = str_split(lines, "\n")[-c(1:5)]
    
    pvals = sapply(lines, function(ll) {
      parts = str_split(ll, " ")
      as.numeric(parts[[1]][length(parts[[1]])])
    })
    
    names(pvals) = sapply(lines, function(ll) {
      parts = str_split(ll, " ")
      parts[[1]][[1]]
    })

    HMAGMA.Pvals[[d]][names(pvals), cond] = pvals 
  }
}


# readr::write_rds(HMAGMA.Pvals, file.path(results.path, "HMAGMA_results_raw.rds"))


```

# Or load preprocessed results
```{r}
HMAGMA.Pvals = readr::read_rds(file.path(results.path, "HMAGMA_results_raw.rds"))

```



## Export selected results
```{r}
# selected.traits = c("hmagmaAdultBrain__sz3", "hmagmaAdultBrain__bip2", "hmagmaAdultBrain__mdd_without_23andMe", "hmagmaAdultBrain__alz2noapoe")
# 
# trait.labels = c("Schizophrenia (SZ)", "Bipolar (BP)", "Depression (MDD)", "Alzheimer (AD)")

  selected.traits = c("hmagmaAdultBrain__sz3", "hmagmaAdultBrain__bip2", "hmagmaAdultBrain__asd", "hmagmaAdultBrain__adhd", "hmagmaAdultBrain__mdd_without_23andMe", "hmagmaAdultBrain__alz2noapoe")

trait.labels = c("Schizophrenia (SZ)", "Bipolar (BP)","Autism (ASD)", "ADHD", "Depression (MDD)", "Alzheimer (AD)")

#   selected.traits = c("hmagmaAdultBrain__sz3", "hmagmaAdultBrain__bip2", "hmagmaAdultBrain__adhd", "hmagmaAdultBrain__alz2noapoe")
# 
# trait.labels = c("Schizophrenia (SZ)", "Bipolar (BP)","ADHD", "Alzheimer (AD)")

# X = -log10(HMAGMA.Pvals$UpAndDown)
# Heatmap(X)


X = HMAGMA.Pvals$UpAndDown[, selected.traits]
cts = intersect(color.df$celltype, rownames(X))
cts = cts[grep("^In|^Ex", cts)]
# 
# X = X[, grep("Adult", colnames(X))]
X = X[cts, ]
# X = matrix(p.adjust(as.numeric(X), "fdr"), nrow = length(cts))
X = apply(X, 2, function(p) p.adjust(p, "fdr"))
rownames(X) = intersect(names(colors), rownames(X))
X = -log10(X)
# X[X < -log10(0.05)] = 0
colnames(X) = trait.labels
rownames(X) = cts

PurPal = colorRampPalette(RColorBrewer::brewer.pal(9, "Purples"))(200)
PurPal = c(rep(PurPal[1], length(PurPal)*(sum(X < 1) / length(X))), PurPal)

require(corrplot)
pdf(file.path(figures.folder, "HMAGMA_adultBrain_UpAndDown_corrplot_logumi_baseline_full_filter_extended_cell_filtering_localFDR_neuro.pdf"), width =7, height = 7)
corrplot(X, is.corr = F, method = "pie", col = PurPal, cl.lim = c(0, 5), cl.length = 5, outline = T, sig.level = 0.05, p.mat = 10^(-X), insig = "blank", tl.col = "black")
dev.off()


```



